---
title: "Intralocus sexual conflict in the DGRP"
author: "Thomas Keaney and Luke Holman"
date: "25/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = FALSE)
```

## Load packages and the data

First load the packages

```{r load packages}

library(tidyverse)
library(googlesheets4) # for reading google sheets
library(pheatmap) # for heatmaps
library(RColorBrewer) # for custom colours
library(rcartocolor) # cool colours
library(wesanderson) # cool colours
library(kableExtra) # for scrolling tables
library(pander) # more tables
library(patchwork) # for cool plots
library(brms) # for bayesian models
library(tidybayes) # for more bayesian things
library(shiny) # for interactive app
library(DT) # for interactive tables
library(broom) # convert results of functions into tables
library(ggExtra) # for marginal plots
library(ggtips) # for hovering tooltips in R shiny
library(ggtext) # for markdown features in ggplot
library(job) # to run things in the background
```

Now load the data. Correlation analysis is not robust when sample sizes are small, so we only include traits that have been measured in 80+ lines. 

```{r}

DGRP_data <- read_csv("data/all.dgrp.phenos_unscaled.csv")

# fix the typo (temporary - keep until we reload the data from google sheets)

DGRP_data$Trait <-recode(DGRP_data$Trait, gustatory.reponse.ethanol.m = "gustatory.response.ethanol.m")

# Now trim the dataset so that it only includes traits that have been measured in 80+ lines - we are left with 905 traits from an original 1119

DGRP_subset <- DGRP_data %>% 
  group_by(Trait) %>% 
  filter(length(line) > 80) %>% 
  ungroup()

# Add a study ID variable 

DGRP_subset <- 
  DGRP_subset %>% 
  group_by(Reference) %>% 
  mutate(study_ID = cur_group_id()) %>% 
  arrange(study_ID) %>% 
  ungroup()

# Assign each trait to a category based on whether it was measured in 1) both sexes separately, 2) just females, 3) just males or 4) both sexes together. We can use a shortcut and use the study IDs as each study tends to be consistent with their measurement approach (with 4 exceptions).

# This is dangerous coding. Using the ID number rather than the name can easily lead to an error. Consider changing.

DGRP_subset <-
  DGRP_subset %>% 
  mutate(rmf_applicable = if_else(study_ID == "3" | study_ID == "5" | study_ID == "7" | study_ID == "8" | study_ID == "9" | study_ID == "11" | study_ID == "13" | study_ID == "19" | study_ID == "20" | study_ID == "21" | study_ID == "23" | study_ID == "24" | study_ID == "26" | study_ID == "27" | study_ID == "28" | study_ID == "34" | study_ID == "37" | study_ID == "40" | study_ID == "43" | study_ID == "48" | study_ID == "49" | study_ID == "54" | study_ID == "55" | study_ID == "57" | study_ID == "64" | study_ID == "65" | study_ID == "71" | study_ID == "72" | study_ID == "73" | study_ID == "75"| study_ID == "76", "YES", "NO"))

# Above, we identified the studies that measured traits in the sexes separately. However, some studies measure multiple traits, some of which they only measure in one sex or both (e.g. larval traits are often measured with the sexes pooled). Below we remove the specific traits that are not measured in both sexes separately from our already trimmed data  

DGRP_rmf <- 
  DGRP_subset %>% 
  filter(rmf_applicable == "YES") %>% 
  filter(Trait != "anoxia.tolerance.larval",
         Trait != "CHC.11 & 9-Me-C23.f", 
         Trait != "CHC.13- & 11-Me-C25.f",
         Trait != "CHC.2-Me-C26.m", 
         Trait !="CHC.2-Me-C27.m", 
         Trait !="CHC.2-Me-C28 & 7,11-C29:2.f",
         Trait != "CHC.2-Me-C28.m", 
         Trait != "CHC.2-Me-C29 & 8,12-C30:2.f",
         Trait != "CHC.2-Me-C29.m", 
         Trait != "CHC.2-Me-C30.f",
         Trait != "CHC.3-Me-C23.m", 
         Trait != "CHC.3-Me-C25 & 9-C26:1.f",
         Trait != "CHC.3-Me-C25.m",
         Trait != "CHC.3-Me-C27 & 6,10-C28:2 & 9-C28:1.f",
         Trait != "CHC.5,9-C25:2 & 9-C25:1.f", 
         Trait != "CHC.5,9-C27:2 & 9-C27:1.f",
         Trait != "CHC.5-C27:1.f", 
         Trait != "CHC.5-Me-C25.f",
         Trait != "CHC.6,10-C26:2.f", 
         Trait != "CHC.6-C26:1 & i.m",
         Trait != "CHC.7,11-C23:2.f", 
         Trait != "CHC.7,11-C24:2.f", 
         Trait != "CHC.7,11-C25:2.f",
         Trait != "CHC.7,11-C26:2.f", 
         Trait != "CHC.7,11-C27:2 & 2-Me-C26.f",
         Trait != "CHC.7,11-C28:2 & 2-Me-C27.f", 
         Trait != "CHC.7,11-C31:2.f",
         Trait != "CHC.7-C24:1.m",
         Trait != "CHC.8,12-C26:2.f",
         Trait != "CHC.8,12-C28:2.f",
         Trait != "CHC.8-C25:1.f",
         Trait != "CHC.9,13-C25:2.f", 
         Trait != "CHC.9,13-C27:2.f",
         Trait != "CHC.9-C25:1.m", 
         Trait != "CHC.9-C27:1.m",
         Trait != "CHC.9-C29:1 & 5,9-C29:2.f", 
         Trait != "CHC.n-C25.m",
         Trait != "CHC.NI.20.f", 
         Trait != "CHC.NI.33.f", 
         Trait != "CHC.NI.34.f",
         Trait != "CHC.NI.7.m", 
         Trait != "CHC.x-C22:1a.m",
         Trait != "CHC.x-C22:1b.m", 
         Trait != "CHC.x-C23:1.f",
         Trait != "kallithea.viral.load.f")

# Make a female trait dataset

DGRP_rf_unscaled <-
  DGRP_rmf %>% filter(Sex == "Female") %>% 
  rename(Female_trait_value = trait_value) %>% 
  select(-c(Sex, `Trait description`, study_ID, rmf_applicable)) %>% 
  mutate(Trait = gsub('.{2}$', '', Trait)) # remove the final two characters from the name of each trait: .f

DGRP_rf <-
  DGRP_rf_unscaled %>% 
  group_by(Trait) %>% # group the data by trait so that scaling works on a per trait basis
  mutate(Female_trait_value = as.numeric(scale(Female_trait_value))) %>% # scale the data for each trait 
  ungroup() 
 
# Make a male trait dataset          

DGRP_rm_unscaled <-
  DGRP_rmf %>% filter(Sex == "Male") %>% 
  rename(Male_trait_value = trait_value) %>% 
  select(-c(Sex, `Trait description`, study_ID, rmf_applicable)) %>%
  mutate(Trait = gsub('.{2}$', '', Trait)) # remove the final two characters from the name of each trait: .m

DGRP_rm <-
  DGRP_rm_unscaled %>% 
  group_by(Trait) %>% # group the data by trait so that scaling works on a per trait basis
  mutate(Male_trait_value = as.numeric(scale(Male_trait_value))) %>% # scale the data for each trait
  ungroup() 

# Join the two together      

rfm_unscaled_data <- full_join(DGRP_rf_unscaled, DGRP_rm_unscaled)

rfm_data <- full_join(DGRP_rf, DGRP_rm)

# Now return to our larger tibble: DGRP_subset. Change the format of the data and scale it.

# This next section changes the format of the data. Each row represents a single line, while each column represents a trait and contains the lines trait mean. Scaling is also applied to each trait value, so that each trait has a mean of 0 and an SD of 1. This makes the variation found for each trait broadly comparable with all other traits.

DGRP_subset_scaled <- 
  split(DGRP_subset, DGRP_subset$Trait) %>% # split divides the data into trait groups
  lapply(function(x){
    trait_name <- x$Trait[1]
    x <- x[,c(1,4)]
    names(x) <- c("line", "value")
    x <- x %>% 
      group_by(line) %>% 
      summarise(mean = mean(value))
    names(x) <- c("line", trait_name)
    x}) %>% 
  purrr::reduce(full_join, by = "line", match = "first") %>% 
  mutate_at(vars(-("line")), ~ as.numeric(scale(.x))) # this is the scaling step

```

$~$

## Calculate the intersex genetic correlation for fitness r^A^~*w fm*~ 

```{r}

# Note that because we are using standarised data, the intercept will always = 0 and the sigma = 1 (we have no predictors so sigma just predicts the population SD). We make the intercept 0 because we have standardised the mean for every trait to = 0. We fix sigma to 0, which makes our estimate more precise. We make sigma = 0 rather than 1 because brms by default actually models the log of sigma. log(1) = 1, so lets set sigma to 0.

rfm_fitness_model <-
  brm(data = rfm_data %>% filter(Trait == "fitness.early"),
      family = gaussian,
      bf(mvbind(Female_trait_value, Male_trait_value) ~ 0, sigma ~ 0) + set_rescor(TRUE),
      prior(lkj(2), class = rescor),
      chains = 4, cores = 4, 
      seed = 1, file = "fits/rfm_fitness.model")   

posterior_summary(rfm_fitness_model) %>% 
  kable()
```

$~$

## Plot r^A^~*w fm*~

```{r}

DGRP_no_line <- 
  DGRP_subset_scaled %>% 
  select(-line)

pos_colour <- wes_palette("Zissou1")[4]
neg_colour <- wes_palette("Zissou1")[2]

fitness_regression_plot <- function(a, b, c, d) {
  ggplot(data = a, aes(x = b, y = c)) +
    #geom_hex(bins = 20) +
    #scale_fill_carto_c(palette = "Temps") +
    geom_point(shape = 21, alpha = 0.75, size = 3, fill = d, show.legend = FALSE) +
    geom_smooth(method = 'lm', color='black') +
    geom_hline(yintercept = 0, linetype = 2) +
    geom_vline(xintercept = 0, linetype = 2) +
    #coord_cartesian(xlim = c(-3, 3), ylim = c(-3, 3)) +
    theme_bw() +
    theme(panel.border= element_blank(),
          axis.line=element_line(),
          text = element_text(size=14),
          panel.grid.major.x = element_blank(),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.minor.x = element_blank(),
          axis.title.x = element_text(size = 12),
          axis.title.y = element_text(size = 12))
}


rfm_fitness_plot <-
  fitness_regression_plot(DGRP_no_line, 
                          DGRP_no_line$fitness.early.f,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "Female fitness", y = "Male fitness") +
  annotate(
    "text", label = "residual correlation = 0.21 (95% CIs: 0.05 to 0.36)",
    x = -2.4, y = 2.55, size = 4, colour = "black"
  )

ggMarginal(rfm_fitness_plot,
      type = 'histogram',
      margins = 'both',
      size = 5,
      colour = wes_palette("Zissou1")[1],
      fill = wes_palette("Zissou1")[2]
    )

```

**Figure 1**: The intersex genetic correlation for fitness r^A^~*w fm*~ across the DGRP. The orange points each represent a DGRP line, and where it falls on the female-male fitness continuum. Data points are standardised to have a mean of 0 and a standard deviation of 1. This means that lines that have negative values have lower than average fitness, while lines that have positive values have above average fitness. The dotted lines separate the plot into four quadrants: lines that fall within the top left quadrant have above average male fitness, but below average female fitness, lines in the top right have above average fitness for both sexes, lines in the bottom left have below average fitness in both sexes, and lines in the bottom right have below average fitness in males, but above average fitness in females.

$~$

## Calculating r^A^~*z fm*~ for all traits

$~$

The traits we consider in this analysis fall into four categories. 1) They have been measured in the two sexes separately, 2) they have only been measured females, 3) they have only been measured in males and 4) they have been measured across the two sexes. For those traits that fall into category 1, we can calculate r^A^~*fm*~.

We use bayesian models to find the residual correlation between traits expressed in males and females. Each line represents a different genotype, and we have measures for both sexes per line. This correlation therefore corresponds to the additive intersex genetic correlation for each trait.

A problem here is that my current approach assumes that each trait is normally distributed, which is clearly false. This is the same problem that researchers face with the spearman correlation coefficient, and the main reason we have used the pearson method in earlier analyses. To solve this problem, I can categorise each trait into a 'distribution family' then run the models in subsets. Using this bayesian approach is preferred because it gives us measures of uncertainty which you don't get from the pearson frequentist approach. **Currently not done**

```{r, fig.height=28, fig.width=9}

# find all the traits in our dataset

check <- unique(rfm_data$Trait)

# this is our model outline - we can update this with new filtered data to calculate rfm for every trait. We do this because we don't want brms to recompile every time. With the model already compiled we can simply update, which saves a load of time.

rfm_model <-
  brm(data = rfm_data %>% filter(Trait == "gustatory.response.ethanol"),
      family = gaussian,
      bf(mvbind(Female_trait_value, Male_trait_value) ~ 0, sigma ~ 0) + set_rescor(TRUE),
      prior(lkj(2), class = rescor),
      chains = 4, cores = 4, 
      seed = 1)   

# sample from the posterior - this will create a tibble with two columns - Trait and rfm

post <- posterior_samples(rfm_model) %>% 
  rename(rfm = rescor__Femaletraitvalue__Maletraitvalue) %>% 
  mutate(Trait = "olfactory.response.1-hexanol") %>% 
  select(Trait, rfm) %>% 
  as_tibble()

# make a function to update the model and the posterior sample output with your desired trait

rmf_calculator <- function(selected_trait){
  
  data <- rfm_data %>% filter(Trait == selected_trait)
  
  model <- update(
    rfm_model, newdata = data,
    chains = 4, cores = 4, 
    seed = 1)
  
  post <- 
    posterior_samples(model) %>% 
    rename(rfm = rescor__Femaletraitvalue__Maletraitvalue) %>% 
    mutate(Trait = selected_trait) %>% 
    select(Trait, rfm) %>% 
    as_tibble() #%>% 
    rbind(post) # we add this to the bottom of the previous post object: it gains progressively more rows as traits are added
  
  post
}

# create a list of all trait to iterate our function over

trait_list <- unique(rfm_data$Trait)

# run the function

if(!file.exists("data/rfm.complete.csv")){
  rfm_complete <- map_dfr(trait_list, rmf_calculator)
  write_csv(rfm_complete, file = "data/rfm.complete.csv")
} else rfm_complete <- read_csv("data/rfm.complete.csv")



# now plot rfm for all traits

forest_plot <-
  rfm_complete %>% 
  group_by(Trait) %>% 
  mutate(avg_rfm = median(rfm)) %>%
  ggplot(aes(rfm, fct_reorder(Trait, avg_rfm))) +
  stat_interval(.width = c(0.05, 0.66, 0.95), 
                height = 3, show.legend = F) +
  rcartocolor::scale_color_carto_d(palette = "Peach") +
  #stat_pointinterval(aes(), .width = c(0.66, 0.95)) +
  coord_cartesian(xlim = c(-0.5, 1)) +
  ylab("Trait") +
  xlab("Estimated _r_^A z fm") +
  theme_bw() + 
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        axis.title.x = ggtext::element_markdown())

forest_plot

```

**Figure X**: The intersex additive genetic correlation for traits that have been measured in both sexes in > 80 DGRP lines. A value of 0 indicates that there is no intersex genetic correlation, while a value of 1 means that additive genetic effects in one sex affect the other one identically. Data are means (depicted as 5% uncertainty intervals) with 66% and 95% uncertainty intervals.

$~$

This figure is cluttered, mainly because there are so many microbiome traits. Below I attempt to visualise the data without these microbiome traits. This code is also useful because we can tag each trait by its guild.

**It has problems**

```{r, fig.height= 16, fig.width=9, eval=FALSE}

# I don't really get what left_join is doing here. Why do we have 186 million rows! I just want the trait guild and reference columns added to rfm_complete, grouping by trait. I would like to retain all of the rows in rfm_complete (which has many more rows than rfm_data).

# the distinct() hack I use here is not great. It removes values draws from the posterior that are the same by chance, instead of because I have accidentally duplicated them with my full_join. This removes ~50,000 of the ~ 1 million rows that should be retained. 

rfm_complete_guilds <- 
  full_join(rfm_complete, rfm_data %>% 
              select(-line)) #%>% 
  distinct()

# now plot rfm in guild groups

forest_plot_2 <-
  rfm_complete_guilds %>%
  filter(`Trait guild` != "Microbiome") %>% 
  group_by(Trait) %>% 
  mutate(avg_rfm = median(rfm)) %>%
  ggplot(aes(rfm, fct_reorder(Trait, avg_rfm))) +
  stat_interval(.width = c(0.05, 0.66, 0.95), 
                height = 3, show.legend = F) +
  rcartocolor::scale_color_carto_d(palette = "Peach") +
  #stat_pointinterval(aes(), .width = c(0.66, 0.95)) +
  coord_cartesian(xlim = c(-0.5, 1)) +
  ylab("Trait") +
  xlab("Estimated _r_^A z fm") +
  theme_bw() + 
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        axis.title.x = ggtext::element_markdown())

```

## Calculating sexual dimorphism for all traits

$~$

Following Bijl and Mank (2021) we can calculate sexual dimorphism as: $SD = \frac{\tilde{x}_{larger.sex}}{\tilde{x}_{smaller.sex}} - 1$

To do this we need:

- trait values that are standardised with the sexes pooled

- the data in a different format: male and female trait values in a single column, with a new column identifying the sex of the individual

- then we do it bayes style, where the means come as a distribution

- the hard part here is finding an efficient way to divide the larger trait mean by the smaller one, because this will often flip between traits 

- once again we have issues with normality

```{r}
d1 <- 
  rfm_unscaled_data %>% 
  select(-Male_trait_value) %>% 
  mutate(Sex = "Female") %>% 
  rename(Trait_value = Female_trait_value)

d2 <-
  rfm_unscaled_data %>% 
  select(-Female_trait_value) %>% 
  mutate(Sex = "Male") %>% 
  rename(Trait_value = Male_trait_value)

dimorphism_data <- rbind(d1, d2) %>% 
  group_by(Trait) %>% # group the data by trait so that scaling works on a per trait basis
  mutate(Trait_value = as.numeric(scale(Trait_value))) %>% # scale the data for each trait
  ungroup() 

# now we model dimorphism score to find an average value for each trait 

# no - should be un-scaled and done as above...

dimorphism_model <-
  brm(data = dimorphism_data,
      family = gaussian,
      Trait_value ~ 1 + Sex + (Sex|Trait),
      prior = c(prior(normal(0, 3), class = b),
                prior(normal(0, 5), class = Intercept),
                prior(cauchy(0, 2), class = sd)),
      chains = 4, cores = 4, warmup = 3000, iter = 10000, 
      seed = 1, file = "fits/dimorphism_model") 

#job::job({
dimorphism_model_2 <-
  brm(data = dimorphism_data,
      family = gaussian,
      Trait_value ~ 1 + (Sex|Trait),
      prior = c(prior(normal(0, 5), class = Intercept),
                prior(cauchy(0, 2), class = sd)),
      chains = 4, cores = 4, warmup = 3000, iter = 8000, 
      seed = 1, file = "fits/dimorphism_model_2") 
#})

dimorphism_model_2

new_data <-
  expand_grid(Trait = unique(dimorphism_data$Trait),
              Sex = unique(dimorphism_data$Sex))

dimorphism_predictions <- fitted(dimorphism_model_2, new_data) %>% 
  as_tibble()

dimorphism_predictions <- cbind(new_data, dimorphism_predictions)

```

$~$

## Calculating *S*: the selection differential

$~$

We can calculate the selection differential associated with each phenotype using the **Robertson Price identity**, which states that the selection differential is equivalent to the covariance between phenotype and relative fitness

$S = \sigma(w_i, z_i)$

or

$S = cov(relative.fitness, standarised.phenotype)$

To do this, I currently use the pearson correlation coefficient. I would like to use bayesian analysis instead, so that we get uncertainty estimates on the selection differentials.

We can also calculate the broad sense heritability associated with each phenotype... if we could calculate the narrow sense heritability we could use the **Breeder's equation** to calculate the response to selection for each phenotype (seems unlikely).

**Table 1:** Selection coefficients ($S$) for each phenotype, when expressed in each sex. The antagonism index column shows how different selection on the trait is in the two sexes. Positive values indicate that the trait has a more positive selection differential in females than it does in males, a value of 0 indicates that the selection differential is the same in each sex, and negative values indicate a more positive selection differential in males compared to females. 

```{r, fig.height=75, fig.width=16}

# Calculate female selection differentials - identify all traits that don't end in .m

Female_selection_data <- 
  DGRP_subset_scaled %>% 
  select(!ends_with(".m"))

# Calculate the correlation between female fitness and all these traits - maybe we go univariate here...

# this is our model outline - we can update this with new filtered data to calculate rfm for every trait. We do this because we don't want brms to recompile every time. With the model already compiled we can simply update, which saves a load of time.

data <- Female_selection_data %>% 
  select("fitness.early.f", "Acetobacter.aceti.f") %>% 
  rename(V1 = 1, V2 = 2)

s_female_model<-
  brm(data = data,
      family = gaussian,
      V1 ~ 0 + V2,
      prior = c(prior(normal(0, 1), class = b),
                prior(normal(0, 1), class = sigma)),
      chains = 4, cores = 4, 
      seed = 1)

# make a function to update the model and the posterior sample output with your desired trait

s_female_calculator <- function(selected_trait){
  
  data <- Female_selection_data %>% 
    select("fitness.early.f", selected_trait) %>% 
    rename(V1 = 1, V2 = 2)
  
  model <- update(
    s_female_model,
    newdata = data,
    chains = 4, cores = 4, 
    seed = 1)
  
  post <- 
    posterior_samples(model) %>% 
    rename(S_female = 1) %>% 
    mutate(Trait = selected_trait) %>% 
    select(Trait, S_female) %>% 
    as_tibble() #%>% 
  rbind(post) # we add this to the bottom of the previous post object: it gains progressively more rows as traits are added
  
  post
}

# create a list of all trait to iterate our function over

trait_list_female <- colnames(Female_selection_data %>% select(-fitness.early.f))

# run the function

if(!file.exists("data/S_female.csv")){
  S_female <- map_dfr(trait_list_female, s_female_calculator)
  write_csv(S_female, file = "data/S_female.csv")
} else S_female <- read_csv("data/S_female.csv")

S_female <-
  S_female %>% 
  filter(Trait != "line")

# now plot S_female for all traits

S_female_forest_plot <-
  S_female %>% 
  group_by(Trait) %>% 
  mutate(avg_S = median(S_female)) %>%
  ggplot(aes(S_female, fct_reorder(Trait, avg_S))) +
  stat_interval(.width = c(0.05, 0.66, 0.95), 
                height = 3, show.legend = F) +
  rcartocolor::scale_color_carto_d(palette = "Peach") +
  #stat_pointinterval(aes(), .width = c(0.66, 0.95)) +
  coord_cartesian(xlim = c(-0.5, 1)) +
  ylab("Trait") +
  xlab("Estimated female selection differential") +
  theme_bw() + 
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        axis.title.x = ggtext::element_markdown())


# Calculate male selection differentials - identify all traits that don't end in .f

Male_selection_data <- 
  DGRP_subset_scaled %>% 
  select(!ends_with(".f"))

# Repeat process for males

# make a function to update the model and the posterior sample output with your desired trait

s_male_calculator <- function(selected_trait){
  
  male_data <- Male_selection_data %>% 
    select("fitness.early.m", selected_trait) %>% 
    rename(V1 = 1, V2 = 2)
  
  model <- update(
    s_female_model,
    newdata = male_data,
    chains = 4, cores = 4, 
    seed = 1)
  
  post <- 
    posterior_samples(model) %>% 
    rename(S_male = 1) %>% 
    mutate(Trait = selected_trait) %>% 
    select(Trait, S_male) %>% 
    as_tibble() #%>% 
  rbind(post) # we add this to the bottom of the previous post object: it gains progressively more rows as traits are added
  
  post
}

# create a list of all trait to iterate our function over

trait_list_male <- colnames(Male_selection_data %>% select(-fitness.early.m))

# run the function

if(!file.exists("data/S_male.csv")){
  S_male <- map_dfr(trait_list_male, s_male_calculator)
  write_csv(S_male, file = "data/S_male.csv")
} else S_male <- read_csv("data/S_male.csv")

S_male <-
  S_male %>% 
  filter(Trait != "line")

# now plot S_female for all traits

S_male_forest_plot <-
  S_male %>% 
  group_by(Trait) %>% 
  mutate(avg_S = median(S_male)) %>%
  ggplot(aes(S_male, fct_reorder(Trait, avg_S))) +
  stat_interval(.width = c(0.05, 0.66, 0.95), 
                height = 3, show.legend = F) +
  rcartocolor::scale_color_carto_d(palette = "Peach") +
  #stat_pointinterval(aes(), .width = c(0.66, 0.95)) +
  coord_cartesian(xlim = c(-0.5, 1)) +
  ylab(NULL) +
  xlab("Estimated male selection differential") +
  theme_bw() + 
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        axis.title.x = ggtext::element_markdown())

S_female_forest_plot + S_male_forest_plot

```

**Figure XX**: selection differentials for a) females and b) males

$~$

## Identifying traits under sexually antagonistic selection

$~$

We have found selection differentials acting on traits in females and males. For those traits where we have a measure of both, we can quantify sexually antagonistic selection by finding traits that have selection operating in opposite directions, depending on which sex the trait is expressed in. 

[Innocenti and Morrow (2011)]( https://doi.org/10.1111/j.1558-5646.2010.01021.x) present an index for measuring the intensity for sexually antagonistic selection:

$I = \frac{\beta'_M \beta'_F}{\sqrt{\frac{{\beta'_M}^2 {\beta'_F}^2}{2}}}$

Where $B'_F$ = $S_F$ and $B'_M$ = $S_M$ that we have calculated above.

From **Innocenti and Morrow**:

This index is positive when selection is concordant in the two sexes, negative when antagonistic in the two sexes, and is zero when selection is absent in one sex (note that it will miss conflict that occurs when strong stabilizing selection is present in that sex). Finally it is proportional to the absolute intensity of selection.

Additionally, it has the desirable properties of being symmetrical and normally distributed for a random set of normally distributed β′M,β′F. |I| is also always included in the interval between the absolute values of the selection gradient in the two sexes, and it coincides with them when β′M = β′F. As a potential drawback, it should be noted that such quantity is not defined when β′M = β′F = 0, even though it makes little sense to estimate how concordant or antagonstic directional selection is when it is absent in both sexes.


```{r, fig.height=50, fig.width=6}

S_female_a <-
  S_female %>% 
  filter(str_ends(Trait, ".f")) %>% 
  mutate(Trait = gsub('.{2}$', '', Trait))

S_female_b <-
  S_female %>% 
  filter(!str_ends(Trait, ".f")) 

S_female <- rbind(S_female_a, S_female_b) %>% 
  group_by(Trait) %>% 
  mutate(draw = 1:n())

S_male_a <-
  S_male %>% 
  filter(str_ends(Trait, ".m")) %>% 
  mutate(Trait = gsub('.{2}$', '', Trait))

S_male_b <-
  S_male %>% 
  filter(!str_ends(Trait, ".m"))

S_male <- rbind(S_male_a, S_male_b) %>% 
  group_by(Trait) %>% 
  mutate(draw = 1:n())

# Create the combined dataset

S_complete <- full_join(S_female, S_male)

# Calculate Morrow and Innocenti's concordance of selection index

S_complete <-
  S_complete %>% 
  select(-draw) %>% 
  mutate(SI = S_female * S_male / sqrt(((S_female)^2 + (S_male)^2)/2))

Selection_index_forest_plot <-
  S_complete %>% filter(!is.na(SI)) %>% 
  group_by(Trait) %>% 
  mutate(avg_SI = median(SI)) %>%
  ggplot(aes(SI, fct_reorder(Trait, avg_SI))) +
  stat_interval(.width = c(0.05, 0.66, 0.95), 
                height = 3, show.legend = F) +
  rcartocolor::scale_color_carto_d(palette = "Peach") +
  #stat_pointinterval(aes(), .width = c(0.66, 0.95)) +
  coord_cartesian(xlim = c(-0.5, 1)) +
  ylab(NULL) +
  xlab("Concordance of selection index") +
  theme_bw() + 
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        axis.title.x = ggtext::element_markdown())

Selection_index_forest_plot

```

```{r, eval=FALSE}
# Here's a table showing the correlations

S_complete %>% 
  group_by(Trait) %>% 
  summarise()

my_data_table <- function(df){
  datatable(
    df, rownames=FALSE,
    autoHideNavigation = TRUE,
    extensions = c("Scroller",  "Buttons"),
    options = list(
      dom = 'Bfrtip',
      deferRender=TRUE,
      scrollX=TRUE, scrollY=800,
      scrollCollapse=TRUE,
      buttons =
        list('pageLength', 'colvis', 'csv', list(
          extend = 'pdf',
          pageSize = 'A4',
          orientation = 'landscape',
          filename = 'selection_data')),
      pageLength = 905
    )
  )
}

my_data_table(trait_selection_gradients)
```


```{r, eval=FALSE, include=FALSE}

# We need to know how all of our traits correlate with fitness. Let's make looking at this easier by calculating the correlation matrix with just our two fitness parameters.

fef.matrix <- round(cor(y = DGRP_no_line$fitness.early.f, x = DGRP_no_line, use = "pairwise.complete.obs"), 2)
#flf.matrix <- round(cor(y = DGRP_no_line$female.fitness.late, x = DGRP_no_line, use = "pairwise.complete.obs"), 2)
mef.matrix <- round(cor(y = DGRP_no_line$fitness.early.m, x = DGRP_no_line, use = "pairwise.complete.obs"), 2)
#mlf.matrix <- round(cor(y = DGRP_no_line$male.fitness.late, x = DGRP_no_line, use = "pairwise.complete.obs"), 2)

fitness.matrix <- cbind(fef.matrix, mef.matrix) 
  
colnames(fitness.matrix) <- c("S female", "S male")

# calculate a sexual conflict score for each trait: the `antagonism` index.

trait_selection_gradients <- tibble(Trait = names(DGRP_no_line)) %>% 
  cbind(fitness.matrix %>% as_tibble()) %>% as_tibble() %>% 
  mutate(`antagonism index` = `S female` - `S male`) %>% 
  filter(Trait != "fitness.early.m",
         Trait != "fitness.early.f",
         Trait != "fitness.late.m",
         Trait != "fitness.late.f")

# Here's a table showing the correlations

my_data_table <- function(df){
  datatable(
    df, rownames=FALSE,
    autoHideNavigation = TRUE,
    extensions = c("Scroller",  "Buttons"),
    options = list(
      dom = 'Bfrtip',
      deferRender=TRUE,
      scrollX=TRUE, scrollY=800,
      scrollCollapse=TRUE,
      buttons =
        list('pageLength', 'colvis', 'csv', list(
          extend = 'pdf',
          pageSize = 'A4',
          orientation = 'landscape',
          filename = 'selection_data')),
      pageLength = 905
    )
  )
}

my_data_table(trait_selection_gradients)

```

$~$



```{r, eval=FALSE}

# There are some issues here. It does not make sense to calculate the covariance between e.g. female fitness and a phenotype expressed in males, when we also have a measure of the phenotype in females.

# this is not perfect

ggplot(data = trait_selection_gradients, aes(x = `S female`, y = `S male`, fill = `antagonism index`)) +
  #geom_hex(bins = 20) +
  geom_jitter(shape = 21, alpha = 0.8, size = 4, show.legend = FALSE, width = 0.003, height = 0.003) +
  #geom_point(shape = 21, alpha = 0.9, size = 5, show.legend = FALSE) +
  geom_smooth(method = 'lm', color='black') +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  #coord_cartesian(xlim = c(-3, 3), ylim = c(-3, 3)) +
  scale_fill_carto_c(palette = "Geyser",
                     breaks=c(-1, -0.5, 0, 0.5, 1),
                     limits=c(-1,1)) +
  theme_bw() +
  theme(panel.border= element_blank(),
        axis.line=element_line(),
        text = element_text(size=14),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12))

```

$~$

While this figure comes from relatively rough data, we can see that trait values that are postively selected in females tend also to be positively selected in males. However, there are numerous traits that appear to be under sexually antagonstic selection: that is, trait values that make fit males make low fitness females and vice versa. 

## Female fitness

```{r, fig.width= 15, fig.height= 10}

# I'm trying to find a way to automatically find the strongest correlations with each fitness component 

# Now let's plot the trait correlations that are most positively correlated  

early.female.1 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$glutamic.acid.low.yeast.f,
                          DGRP_no_line$fitness.early.f, pos_colour) +
  labs(x = "Level of glutamic\nacid on low yeast\ndiet", y = "Female fitness")

early.female.2 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$lifetime.fecundity.25C.f	,
                          DGRP_no_line$fitness.early.f, pos_colour) +
  labs(x = "Lifetime fecundity", y = NULL)

early.female.3 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$week1.fecundity.25C.f	,
                          DGRP_no_line$fitness.early.f, pos_colour) +
  labs(x = "Week 1 fecundity\n at 25C", y = NULL)

early.female.4 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$M.anisopliae.resistance.f	,
                          DGRP_no_line$fitness.early.f, pos_colour) +
  labs(x = "Resistance to\nM. anisopliae", y = NULL)

early.female.5 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$acetylcarnitine.low.yeast.f	,
                          DGRP_no_line$fitness.early.f, pos_colour) +
  labs(x = "Level of\nacetylcarnitine on low\nyeast diet", y = NULL)

early.female.6 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$circadian.period.mesa.f		,
                          DGRP_no_line$fitness.early.f, pos_colour) +
  labs(x = "Circadian period length", y = "Female fitness")

early.female.7 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$ornithine.low.yeast.f	,
                          DGRP_no_line$fitness.early.f, pos_colour) +
  labs(x = "Level of ornithine\non low yeast diet", y = NULL)

early.female.8 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$P.aeruginosa.resistance.2017.f	,
                          DGRP_no_line$fitness.early.f, pos_colour) +
  labs(x = "Resistance to\nP. aeruginosa", y = NULL)

early.female.9 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$serine.high.yeast.f	,
                          DGRP_no_line$fitness.early.f, pos_colour) +
  labs(x = "Level of serine\non low yeast diet", y = NULL)

early.female.10 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$week3.fecundity.25C.f	,
                          DGRP_no_line$fitness.early.f, pos_colour) +
  labs(x = "Week 3 fecundity\n at 25C", y = NULL)

# Now lets do the most negative correlations

early.female.11 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$epinephrine.high.yeast.f	,
                          DGRP_no_line$fitness.early.f, neg_colour) +
  labs(x = "Level of epinephrine\non high yeast diet", y = "Female fitness")

early.female.12 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$cGMP.low.yeast.f	,
                          DGRP_no_line$fitness.early.f, neg_colour) +
  labs(x = "Level of cGMP\non low yeast diet", y = NULL)

early.female.13<-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$day.sleep.bout.length.f	 ,
                          DGRP_no_line$fitness.early.f, neg_colour) +
  labs(x = "Day sleep bout\n length", y = NULL)

early.female.14 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$Sphingomonas.wittichii.f	 ,
                          DGRP_no_line$fitness.early.f, neg_colour) +
  coord_cartesian(xlim = c(-3, 5), ylim = c(-3, 3)) +
  labs(x = "S. wittichii load\n(microbiome)", y = NULL)

early.female.15 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$`alpha-amanitin.resistance.0.2`	,
                          DGRP_no_line$fitness.early.f, neg_colour) +
  labs(x = "Resistance to mushroom toxin\n (alpha-amanitin 0.2 ug/g)", y = NULL)

early.female.16 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$sprinter.marathoner.f	,
                          DGRP_no_line$fitness.early.f, neg_colour) +
  labs(x = "'Sprinter' phenotype", y = "Female fitness")

early.female.17 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$`CHC.8-C25:1.f`	,
                          DGRP_no_line$fitness.early.f, neg_colour) +
  labs(x = "Level of the CHC\n 8-C25:1", y = NULL)

early.female.18 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$cold.shock.young.f	,
                          DGRP_no_line$fitness.early.f, neg_colour) +
  labs(x = "Resistance to cold\n shock", y = NULL)

early.female.19 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$proline.low.yeast.f	,
                          DGRP_no_line$fitness.early.f, neg_colour) +
  labs(x = "Level of proline\non low yeast diet", y = NULL)

early.female.20 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$`3-Hydroxykynurenine.high.yeast.f`	,
                          DGRP_no_line$fitness.early.f, neg_colour) +
  labs(x = "Level of\n3-Hydroxykynurenine on\n low yeast diet", y = NULL)

(early.female.1 | early.female.2 | early.female.3 | early.female.4 | early.female.5) /
  (early.female.6 | early.female.7 | early.female.8 | early.female.9 | early.female.10) /
  (early.female.11 | early.female.12 | early.female.13| early.female.14| early.female.15) /
  (early.female.16 | early.female.17 | early.female.18 | early.female.19 | early.female.20)



```

**Figure 2**: The 10 phenotypes with the strongest positive (orange points) and negative (blue points) selection coefficients in females. All traits are standardised to have a mean = 0 and SD = 1.

$~$

## Male fitness

```{r, fig.width= 15, fig.height= 10}

# I'm trying to find a way to automatically find the strongest correlations with each fitness component 

young.male.1 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$locomotor.frequency.f,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "Locomotor activity", y = "Male fitness")

young.male.2 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$P.aeruginosa.resistance.2017.m,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "Resistance to bacterial infection\n (P. aeruginosa)", y = NULL)

young.male.3 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$lifespan.28C.m,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "Lifespan at 28C", y = NULL)

young.male.4 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$RCH.old.m,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "Rapid cold hardening\n in old flies", y = NULL)

young.male.5 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$Enterococcus.hirae.m	,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "E. hirae load\n(microbiome)", y = NULL)

young.male.6 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$M.anisopliae.resistance.m	,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "Resistance to fungal infection\n (M. anisopliae)",
       y = "Male fitness")

young.male.7 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$acetylcholine.low.yeast.f	,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "Level of\nacetylcholine on\n low yeast diet", y = NULL)

young.male.8 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$glucose.low.glucose.diet.m	,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "Level of glucose\non low glucose diet", y = NULL)

young.male.9 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$Streptomyces.sp..m	,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "Streptomyces sp. load\n(microbiome)", y = NULL)

young.male.10 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$day.sleep.bout.length.SD.m	,
                          DGRP_no_line$fitness.early.m, pos_colour) +
  labs(x = "Variation in day\nsleep bout length", y = NULL)

# Now lets do the most negative correlations

young.male.11 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$`CHC.6,10-C26:2.f`	,
                          DGRP_no_line$fitness.early.m, neg_colour) +
  labs(x = "Level of the CHC\n 6, 10-C26:2\n(only produced by females)", y = NULL)

young.male.12 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$P.rettgeri.load.high.glucose.diet.m	 ,
                          DGRP_no_line$fitness.early.m, neg_colour) +
  labs(x = "Bacterial load \n(P. rettgeri) 24hrs post\ninfection", y = NULL)

young.male.13 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$`3-Hydroxybutyric.acid.low.yeast.f`	 ,
                          DGRP_no_line$fitness.early.m, neg_colour) +
  labs(x = "Level of\3-Hydroxybutyric acid on\n low yeast diet", y = NULL)

young.male.14 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$cadaverine.high.yeast.f	,
                          DGRP_no_line$fitness.early.m, neg_colour) +
  labs(x = "Level of\ncadaverine on\n high yeast diet", y = NULL)

young.male.15 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$cold.shock.young.m	,
                          DGRP_no_line$fitness.early.m, neg_colour) +
  labs(x = "Resistance to cold\nshock in young flies", y = "Male fitness")

young.male.16 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$`CHC.5,9-C25:2 & 9-C25:1.f`,
                          DGRP_no_line$fitness.early.m, neg_colour) +
  labs(x = "Level of the CHCs\n5,9-C25:2 & 9-C25:1\n(only produced by females)", y = NULL)

young.male.17 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$pantothenate.low.yeast.f,
                          DGRP_no_line$fitness.early.m, neg_colour) +
  labs(x = "Level of\npantothenate on\n low yeast diet", y = NULL)

young.male.18 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$aminoisobutyrate.high.yeast.f	,
                          DGRP_no_line$fitness.early.m, neg_colour) +
  labs(x = "Level of\naminoisobutyrate on\n high yeast diet", y = NULL)

young.male.19 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$`CHC.8,12-C26:2.f`	,
                          DGRP_no_line$fitness.early.m, neg_colour) +
  labs(x = "Level of the CHC\n8,12-C26:2\n(only produced by females)", y = NULL)

young.male.20 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$myristic.acid.low.yeast.f		,
                          DGRP_no_line$fitness.early.m, neg_colour) +
  labs(x = "Level of myristic acid\n on low yeast diet", y = "Male fitness")

(young.male.1 | young.male.2 | young.male.3 | young.male.4 | young.male.5) /
  (young.male.6 | young.male.7 | young.male.8 | young.male.9 | young.male.10) /
  (young.male.11 | young.male.12 | young.male.13 | young.male.14 | young.male.15) /
  (young.male.16 | young.male.17 | young.male.18 | young.male.19 | young.male.20)

```

**Figure 3**: The 10 phenotypes with the strongest positive (orange points) and negative (blue points) selection coefficients in males. All phenotypes are standardised to have a mean = 0 and SD = 1.

$~$

```{r, eval=FALSE, include=FALSE}

# The relationship between fitness and population dynamics

#One study measured several population dynamics across the DGRP - Gerken, Eller and Morgan (2014) They placed 20 females in a vial with two males of the same DGRP line and measured the number of eggs they laid per day and noted daily female mortality. In the advent of male death they were replaced the following day. Using this data the researchers were able to calculate population growth rates, net reproductive rates and estimated generation time. 

#We can look at how our individual fitness components correlate with these population dynamics. The sex ratio is extremely female biased, so it is likely that some females were sperm limited at some stage of their lives, very unlikely that interlocus sexual conflict affected females to a signficant extent, or that different rates of resource consumption between the sexes had a pronounced effect on population fitness Conversely, female fecundity, food intake and male sperm number are likely to have strong effects on population fitness.

#Here's a reminder of the traits measured in Gerken, Eller and Morgan:

left_join(lines, table) %>% 
  select(-line) %>% 
  filter(Reference == "Gerken, Eller and Morgan (2014) PhD thesis",
         Trait != c("RCH.16C", "RCH.23C", "RCH.25C")) %>%
  arrange(Trait) %>% 
  pander(split.cell = 40, split.table = Inf)
```

```{r, fig.width= 15, fig.height= 15, eval=FALSE, include=FALSE}

#The two traits that are most interesting to us are net reproductive rate, which is the lifetime offspring production of the 20 females, and population growth rate, which is the average rate of population increase per day over the entire lives of the 20 females (only the offspring of the 20 original females are counted in this measure). It should be noted that both of these measures are calculated using the same inputs: eggs laid per day and the surviving proportion of females.  

#First lets look at net reproductive output


# First make plots for 16C

early.female.repro.rate <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$female.fitness.early,
                          DGRP_no_line$net.reproductive.rate.16C, pos_colour) +
  labs(x = "Young female fitness", y = "Net reproductive rate\n at 16C")

late.female.repro.rate <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$female.fitness.late,
                          DGRP_no_line$net.reproductive.rate.16C, pos_colour) +
  labs(x = "Old female fitness")

early.male.repro.rate <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$male.fitness.early,
                          DGRP_no_line$net.reproductive.rate.16C, neg_colour) +
  labs(x = "Young male fitness", y = "Net reproductive rate\n at 16C")

late.male.repro.rate <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$male.fitness.late,
                          DGRP_no_line$net.reproductive.rate.16C, pos_colour) +
  labs(x = "Old male fitness") 

# Now plot at 23C

early.female.repro.rate.23 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$female.fitness.early,
                          DGRP_no_line$net.reproductive.rate.23C, pos_colour) +
  labs(x = "Young female fitness", y = "Net reproductive rate\n at 23C")

late.female.repro.rate.23 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$female.fitness.late,
                          DGRP_no_line$net.reproductive.rate.23C, pos_colour) +
  labs(x = "Old female fitness")

early.male.repro.rate.23 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$male.fitness.early,
                          DGRP_no_line$net.reproductive.rate.23C, neg_colour) +
  labs(x = "Young male fitness", y = "Net reproductive rate\n at 23C")

late.male.repro.rate.23 <-
  fitness_regression_plot(DGRP_no_line, DGRP_no_line$male.fitness.late,
                          DGRP_no_line$net.reproductive.rate.23C, pos_colour) +
  labs(x = "Old male fitness") 
 
(early.female.repro.rate | late.female.repro.rate) /
(early.male.repro.rate | late.male.repro.rate) / 
(early.female.repro.rate.23 | late.female.repro.rate.23) /
(early.male.repro.rate.23 | late.male.repro.rate.23) 


#**Figure 6**: Correlations between female and male fitness with net reproductive output. Each point represents the standardised mean for a DGRP line. Yellow points indicate a positive genetic correlation between fitness and reproductive output, while blue points indicate a negative genetic correlation. 

```

```{r, fig.width= 15, fig.height= 15, eval=FALSE, include=FALSE}

#Now lets take a look at $\lambda$, the population growth rate. 

# First make plots for 16C

early.female.pop.growth <-
fitness_regression_plot(DGRP_no_line, DGRP_no_line$female.fitness.early,
                          DGRP_no_line$pop.growth.rate.16C, pos_colour) +
  labs(x = "Young female fitness", y = "Population growth rate\n at 16C")

late.female.pop.growth <-
fitness_regression_plot(DGRP_no_line, DGRP_no_line$female.fitness.late,
                          DGRP_no_line$pop.growth.rate.16C, pos_colour) +
  labs(x = "Old female fitness")

early.male.pop.growth <-
fitness_regression_plot(DGRP_no_line, DGRP_no_line$male.fitness.early,
                          DGRP_no_line$pop.growth.rate.16C, pos_colour) +
  labs(x = "Young male fitness", y = "Population growth rate\n at 16C")

late.male.pop.growth <-
fitness_regression_plot(DGRP_no_line, DGRP_no_line$male.fitness.late,
                          DGRP_no_line$pop.growth.rate.16C, pos_colour) +
  labs(x = "Old male fitness")

### Now plot for 23C

early.female.pop.growth.23 <-
fitness_regression_plot(DGRP_no_line, DGRP_no_line$female.fitness.early,
                          DGRP_no_line$pop.growth.rate.23C, pos_colour) +
  labs(x = "Young female fitness", y = "Population growth rate\n at 23C")

late.female.pop.growth.23 <-
fitness_regression_plot(DGRP_no_line, DGRP_no_line$female.fitness.late,
                          DGRP_no_line$pop.growth.rate.23C, pos_colour) +
  labs(x = "Old female fitness")

early.male.pop.growth.23 <-
fitness_regression_plot(DGRP_no_line, DGRP_no_line$male.fitness.early,
                          DGRP_no_line$pop.growth.rate.23C, neg_colour) +
  labs(x = "Young male fitness", y = "Population growth rate\n at 23C")

late.male.pop.growth.23 <-
fitness_regression_plot(DGRP_no_line, DGRP_no_line$male.fitness.late,
                          DGRP_no_line$pop.growth.rate.23C, pos_colour) +
  labs(x = "Old male fitness")

(early.female.pop.growth | late.female.pop.growth) /
(early.male.pop.growth | late.male.pop.growth) / 
(early.female.pop.growth.23 | late.female.pop.growth.23) /
(early.male.pop.growth.23 | late.male.pop.growth.23) 

#**Figure 7**: Correlations between female and male fitness with population growth rate. Each point represents the standardised mean for a DGRP line. Yellow points indicate a positive genetic correlation between fitness and reproductive output, while blue points indicate a negative genetic correlation. 

```
