---
title: "The response to selection, sexual antagonism and sexual concordance in the DGRP"
author: "Thomas Keaney and Luke Holman"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, cache = FALSE)
```

# Load packages and the data

$~$

First load the packages and build helper functions

```{r load packages}

library(tidyverse) # for tidy coding
library(MetBrewer) # for many nice colour palettes
library(rcartocolor) # more cool colours
library(kableExtra) # for scrolling tables
library(DT) # for interactive tables
library(patchwork) # to join multiple plots nicely
library(brms) # for bayesian models
library(tidybayes) # for more bayesian things
library(bayestestR) # for the pd metric 
library(broom) # convert results of functions into tables
library(ggtext) # for markdown features in ggplot
library(ggrepel) # for plot labels in ggplot
library(ggnewscale) # to reset scales in ggplot 
library(bigsnpr) 

# Create a function to build HTML searchable tables

my_data_table <- function(df){
  datatable(
    df, rownames=FALSE,
    autoHideNavigation = TRUE,
    extensions = c("Scroller",  "Buttons"),
    options = list(
      autoWidth = TRUE,
      dom = 'Bfrtip',
      deferRender=TRUE,
      scrollX=TRUE, scrollY=1000,
      scrollCollapse=TRUE,
      buttons =
        list('pageLength', 'colvis', 'csv', list(
          extend = 'pdf',
          pageSize = 'A4',
          orientation = 'landscape',
          filename = 'Trait_data')),
      pageLength = 100
    )
  )
}

```

$~$

# Organismal-level trait analysis

## Selecting data appropriate for analysis

We conducted a near-exhaustive search of the literature until January 2022, to obtain line mean estimates and associated meta-data for quantitative traits that have been measured in the DGRP. We did not include data collected for traits that had been measured in heterozygous combinations of multiple DGRP lines. In total, we identified 125 studies that reported line means or raw data for 2115 phenotypic traits. To ready the data for analysis, we grouped trait values by trait and sex and standardised the data to have mean = 0 and sd = 1. Sex-specific standardised line means for each trait were then combined with their standardised fitness estimates, obtained from [Wong and Holman (2023)](). We also include some helpful metadata for downstream analysis. We then pruned the dataset to only include traits that have been measured in single-sex cohorts, in 80 or more lines. We also removed traits included in two large datasets on the microbiome and metabolome (Everett _et al_. 2020 and Jin _et al_. 2020).

```{r}

# load in the data, note that traits have already been standardised

DGRP_data <- 
  left_join(
    read_csv("data/input/all.dgrp.phenos_scaled.csv") %>% 
      mutate(line = as.factor(line)),
    read_csv("data/input/meta_data_for_all_traits.csv") %>%
      group_by(Reference) %>% 
      mutate(study_ID = as.factor(cur_group_id()),
             Pooled = if_else(Sex == "Pooled", "Yes", "No"))) %>% 
  left_join(read_rds("data/input/trait_names.rds"))
  
  
# Apply the selection criteria with the filter() function, then add a column each for female and male fitness

trait_data <-
  DGRP_data %>%  
  filter(!str_detect(Trait, "fitness"),
         `# lines measured` >= 80 &
           Pooled != "Yes" & 
           Reference != "Jin et al (2020) PLOS Genetics" & 
           Reference != "Everett et al (2020) Genome Research" &
           !str_detect(Trait, "dopamine.response.to.paraquat.2021.m")) %>% # data for this trait was entered into the database incorrectly, it was only measured in 10 lines so should not be included in our analysis 
  
  # join the early life fitness data from Wong and Holman
  
  left_join(
    DGRP_data %>%
      filter(str_detect(Trait, "fitness.early.life")) %>% 
      select(line, Trait, trait_value) %>% 
      pivot_wider(names_from = Trait, values_from = trait_value) %>% 
      rename(female_fitness = fitness.early.life.f, male_fitness = fitness.early.life.m))


clean_meta_data <-
  trait_data %>% 
  select(Trait_nice, Trait, Life_stage, `Trait guild`, study_ID, Trait_nice, Reference, `Trait description`) %>% 
  distinct(Trait, .keep_all = TRUE) %>% 
    mutate(Trait_nice = case_when(Trait_nice == "DDT resistance mortality" ~ "DDT susceptibility",
                                  .default = Trait_nice)) # make confusing name simpler

```

$~$

**Table S1**. Traits included for analysis

```{r}
# how many studies did we start with?

table_data <- left_join(
  DGRP_data %>%
    distinct(Trait, .keep_all = TRUE) %>% 
    mutate(Measured_in = case_when(Sex == "Female" ~ "Females",
                                   Sex == "Male" ~ "Males",
                                   Sex == "Pooled" ~ "Mixed sex cohorts")) %>% 
    select(Trait_nice, Trait, Reference, `# lines measured`, Measured_in, `Trait description`),
  
  trait_data %>%
    distinct(Trait) %>% 
    mutate(Included = "Yes")
) %>%
  mutate(Included = if_else(is.na(Included), "No", Included)) %>% 
  select(-Trait) %>% 
  rename(Trait = Trait_nice, `Measured in` = Measured_in) %>% 
  arrange(desc(Included))

# Create a function to build HTML searchable tables

my_data_table(table_data %>% 
               select(Trait, Reference, `# lines measured`, `Measured in`, Included, `Trait description`))
```

$~$

Find the number of traits and studies included in our analysis.

```{r}

# how many studies and traits do we have after filtering?

num_unique_traits <- table_data %>% filter(Included == "Yes") %>% nrow()
# in females
num_unique_traits_f <- table_data %>% filter(Included == "Yes" & `Measured in` == "Females") %>% nrow()
# in males
num_unique_traits_m <- table_data %>% filter(Included == "Yes" & `Measured in` == "Males") %>% nrow()
# how many studies are they measured across in total?
num_unique_studies <- table_data %>% filter(Included == "Yes") %>%  distinct(Reference) %>% nrow()
# in females
num_unique_studies_f <- table_data %>% filter(Included == "Yes" & `Measured in` == "Females") %>% distinct(Reference) %>% nrow()
# in males
num_unique_studies_m <- table_data %>% filter(Included == "Yes" & `Measured in` == "Males") %>% distinct(Reference) %>% nrow()

```

After this selection process, `r num_unique_traits` remain, that were measured across `r num_unique_studies` studies. There are `r num_unique_traits_f` measured in females and `r num_unique_traits_m` in males across `r num_unique_studies_f` and `r num_unique_studies_m` respectively.  

$~$

## Calculating *R*: the response to selection

$~$

We assume that line represents a genetically homogeneous, homozygous genotype, such that the line mean for a particular trait equals the breeding value for that trait for that genotype. We can therefore estimate the genetic variance for a given trait in the DGRP as the variance in line means ([Mackay, 2012](https://www.nature.com/articles/nature10811)).

We can then estimate the response to selection ($R$) for a trait using **Robertson's Secondary Theorem of Natural Selection** (also known as the **Robertson covariance**; Robertson, 1968), which states that $R$ for a given trait is equivalent to the additive genetic covariance between the trait ($A_z$) and relative fitness ($A_w$), such that $R = \sigma(A_w, Az)$. We can estimate this genetic covariance as the correlation in breeding values for traits with relative fitness across DGRP lines.

In species with two sexes and obligate sexual reproduction, half of the response to selection is due to selection on females and half to selection on males (e.g. [Lande, 1980](https://www.jstor.org/stable/2407393)). We can therefore partition the secondary theorem into the response via each sex:

$$R = \frac{\sigma(A_w^F, A_z) + \sigma(A_w^M, A_z)}{2}$$

Furthermore, in quantitative genetics it is common to regard the same trait expressed in males or females (e.g. female body size and male body size) as separate traits. For convenience in the code below, we denote the response to selection of a female trait ($R_F$) or a male trait ($R_M$) as:

$$R_F = \frac{R_{FF} + R_{MF}}{2}$$

$$R_M = \frac{R_{MM} + R_{FM}}{2}$$

Where $R_{FF}$ is the response of a female trait to selection on females, $R_{MF}$ is the response of a female trait to selection on males, $R_{MM}$ is the response of a male trait to selection on males, and $R_{FM}$ is the response of a male trait to selection on females. These four quantities can be found via the partitioned Robertson covariance as follows:

$$R_{FF} = \sigma(A_w^F, A_z^F)$$

$$R_{MF} = \sigma(A_w^M, A_z^F)$$

$$R_{MM} = \sigma(A_w^M, A_z^M)$$

$$R_{FM} = \sigma(A_w^F, A_z^M)$$

Note that even though female traits are not expressed in males and therefore are not under selection in males, selection on males can still generate a response to selection in female traits if there is genetic covariance between these female traits and male traits that are under selection in males (and similarly, male traits can respond to selection on females due to between-sex genetic correlations). We therefore expect $R_{MF}$ and $R_{FM}$ to commonly be non-zero.

$~$

### Build models to calculate $R_{FF}$ & $R_{MF}$

To estimate the covariance between $A_w$ and $A_z$, we fitted bivariate Bayesian linear models using the `brms` package (Bürkner, 2017) for `R version 4.2.2`. For each combination of trait and sex, we used line means for the focal trait and the fitness of the focal sex as the two response variables and fitted an intercept-only Gaussian model. Each model returned a posterior distribution of the residual correlation between trait and fitness, which for data expressed in standard units is equivalent to the covariance.

Build functions to run the models

```{r}

# RFF estimates

female_traits <- trait_data %>% filter(Sex == "Female")

trait_list_female <- unique(female_traits$Trait) # an input to the map_dfr() function that we'll need in a few chunks time

# code the model structure we will use for all traits using one example - `flight.performance.f`. We can then use the update() function to run this model many times, once for each trait measured in females. update() makes this process many times faster, because the model can immediately start sampling, without the need to recompile. 

RFF_model <-
  brm(data = female_traits %>% filter(Trait == "flight.performance.f"),
      family = gaussian,
      bf(mvbind(female_fitness, trait_value) ~ 1) + set_rescor(TRUE),
      prior = c(prior(normal(0, 0.1), class = Intercept, resp = femalefitness),
                prior(normal(0, 0.1), class = Intercept, resp = traitvalue),
                prior(normal(1, 0.1), class = sigma, resp = femalefitness),
                prior(normal(1, 0.1), class = sigma, resp = traitvalue),
                prior(lkj(2), class = rescor)),
      chains = 4, cores = 4, iter = 6000, warmup = 2000,
      seed = 1, file = "fits/RFF_test_model")   

# make a function to update the model and the posterior sample output with the 'selected trait'

RFF_calculator <- function(selected_trait){
  
  data <- female_traits %>% filter(Trait == selected_trait)
  
  model <- update(
    RFF_model, newdata = data,
      chains = 4, cores = 4, iter = 6000, warmup = 2000,
      seed = 1)
  
  posterior <- 
    as_draws_df(model) %>% 
    rename(Response_to_selection_female = rescor__femalefitness__traitvalue) %>% 
    mutate(Trait = selected_trait) %>% 
    select(Trait, Response_to_selection_female) %>% 
    as_tibble()
  
  posterior
}

# RMF estimates

RMF_model <-
  brm(data = female_traits %>% filter(Trait == "flight.performance.f"),
      family = gaussian,
      bf(mvbind(male_fitness, trait_value) ~ 1) + set_rescor(TRUE),
      prior = c(prior(normal(0, 0.1), class = Intercept, resp = malefitness),
                prior(normal(0, 0.1), class = Intercept, resp = traitvalue),
                prior(normal(1, 0.1), class = sigma, resp = malefitness),
                prior(normal(1, 0.1), class = sigma, resp = traitvalue),
                prior(lkj(2), class = rescor)),
      chains = 4, cores = 4, iter = 6000, warmup = 2000,
      seed = 1, file = "fits/RMF_test_model")

# make a function to update the model and the posterior sample output with your desired trait

RMF_calculator <- function(selected_trait){
  
  data <- female_traits %>% filter(Trait == selected_trait)
  
  model <- update(
    RMF_model, newdata = data,
    chains = 4, cores = 4, iter = 6000, warmup = 2000,
    seed = 1)
  
  posterior <- 
    as_draws_df(model) %>% 
    rename(Response_to_selection_male = rescor__malefitness__traitvalue) %>% 
    mutate(Trait = selected_trait) %>% 
    select(Trait, Response_to_selection_male) %>% 
    as_tibble()
  
  posterior
}

```

$~$

### Build models to calculate $R_{FM}$ & $R_{MF}$

$~$

```{r}

# RMM estimates

male_traits <- trait_data %>% filter(Sex == "Male")

trait_list_male <- unique(male_traits$Trait)

RMM_model <-
  brm(data = male_traits %>% filter(Trait == "flight.performance.m"),
      family = gaussian,
      bf(mvbind(male_fitness, trait_value) ~ 1) + set_rescor(TRUE),
      prior = c(prior(normal(0, 0.1), class = Intercept, resp = malefitness),
                prior(normal(0, 0.1), class = Intercept, resp = traitvalue),
                prior(normal(1, 0.1), class = sigma, resp = malefitness),
                prior(normal(1, 0.1), class = sigma, resp = traitvalue),
                prior(lkj(2), class = rescor)),
      chains = 4, cores = 4, iter = 6000, warmup = 2000,
      seed = 1, file = "fits/RMM_test_model")   

# make a function to update the model and the posterior sample output with your desired trait

RMM_calculator <- function(selected_trait){
  
  data <- male_traits %>% filter(Trait == selected_trait)
  
  model <- update(
    RMM_model, newdata = data,
    chains = 4, cores = 4, iter = 6000, warmup = 2000,
    seed = 1)
  
  posterior <- 
    as_draws_df(model) %>% 
    rename(Response_to_selection_male = rescor__malefitness__traitvalue) %>%
    mutate(Trait = selected_trait) %>% 
    select(Trait, Response_to_selection_male) %>% 
    as_tibble()
  
  posterior
}

# RFM estimates

RFM_model <-
  brm(data = male_traits %>% filter(Trait == "flight.performance.m"),
      family = gaussian,
      bf(mvbind(female_fitness, trait_value) ~ 1) + set_rescor(TRUE),
      prior = c(prior(normal(0, 0.1), class = Intercept, resp = femalefitness),
                prior(normal(0, 0.1), class = Intercept, resp = traitvalue),
                prior(normal(1, 0.1), class = sigma, resp = femalefitness),
                prior(normal(1, 0.1), class = sigma, resp = traitvalue),
                prior(lkj(2), class = rescor)),
      chains = 4, cores = 4, iter = 6000, warmup = 2000,
      seed = 1, file = "fits/RFM_test_model")   

# make a function to update the model and the posterior sample output with your desired trait

RFM_calculator <- function(selected_trait){
  
  data <- male_traits %>% filter(Trait == selected_trait)
  
  model <- update(
    RFM_model, newdata = data,
      chains = 4, cores = 4, iter = 6000, warmup = 2000,
      seed = 1)
  
  posterior <- 
    as_draws_df(model) %>% 
    rename(Response_to_selection_female = rescor__femalefitness__traitvalue) %>% 
    mutate(Trait = selected_trait) %>% 
    select(Trait, Response_to_selection_female) %>% 
    as_tibble()
  
  posterior
}
```

$~$

### Run the models for all the traits

Run the models using `RFF_calculator`, `RMF_calculator`, `RMM_calculator` and `RFM_calculator`

```{r}
# run the RFF function

Run_function <- FALSE # change this to TRUE to run the models

if(Run_function){
  RFF <- map_dfr(trait_list_female, RFF_calculator) # map_dfr returns a data frame created by row-binding each output
  write_csv(RFF, file = "data/organismal_level_output/RFF.csv")
} else RFF <- read_csv("data/organismal_level_output/RFF.csv")

# run the RMF function

if(Run_function){
  RMF <- map_dfr(trait_list_female, RMF_calculator) 
  write_csv(RMF, file = "data/organismal_level_output/RMF.csv")
} else RMF <- read_csv("data/organismal_level_output/RMF.csv")

# run the RMM function

if(Run_function){
  RMM <- map_dfr(trait_list_male, RMM_calculator) 
  write_csv(RMM, file = "data/organismal_level_output/RMM.csv")
} else RMM <- read_csv("data/organismal_level_output/RMM.csv")

# run the RFM function

if(Run_function){
  RFM <- map_dfr(trait_list_male, RFM_calculator) 
  write_csv(RFM, file = "data/organismal_level_output/RFM.csv")
} else RFM <- read_csv("data/organismal_level_output/RFM.csv")

```

$~$

### Plot the sex-specific responses to selection

```{r, fig.height=30, fig.width=12}
R_female_traits <-
  bind_cols(RFF, RMF) %>% 
  rename(Trait = Trait...1) %>% 
  mutate(Trait_Sex = "Female") %>% 
  select(-(Trait...3))

p1 <-
  R_female_traits %>% 
  group_by(Trait) %>% 
  mutate(avg_R = median(Response_to_selection_female)) %>%
  ggplot(aes(Response_to_selection_female, fct_reorder(Trait, avg_R))) +
  stat_interval(.width = c(0.05, 0.66, 0.95), 
                height = 1, show.legend = F) +
  rcartocolor::scale_color_carto_d(palette = "Purp") +
  coord_cartesian(xlim = c(-0.5, 0.5)) +
  geom_vline(linetype = 2, xintercept = 0, linewidth = 1) +
  labs(x = "Response to selection in females",
       y = "Trait expressed in females") +
  theme_bw() + 
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        text = element_text(size=14),
        axis.text.y = element_text(size = 8))

p2 <-
  R_female_traits %>% 
  group_by(Trait) %>% 
  mutate(avg_R = median(Response_to_selection_male)) %>%
  ggplot(aes(Response_to_selection_male, fct_reorder(Trait, avg_R))) +
  stat_interval(.width = c(0.05, 0.66, 0.95), 
                height = 1, show.legend = F) +
  rcartocolor::scale_color_carto_d(palette = "Peach") +
  coord_cartesian(xlim = c(-0.5, 0.5)) +
  geom_vline(linetype = 2, xintercept = 0, linewidth = 1) +
  labs(x = "Response to selection in males",
       y = "Trait expressed in females") +
  theme_bw() + 
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        text = element_text(size=14),
        axis.text.y = element_text(size = 8)) 

p1 + p2 +
  plot_annotation(tag_levels = 'a')

```

**Figure S1**. The estimated response to selection in **a** females and **b** males for all traits measured in females. Innermost bands approximate the median, while outer bands show the 66 and 95% credible intervals.

$~$

```{r, fig.height=30, fig.width=12}

R_male_traits <-
  bind_cols(RFM, RMM) %>% 
  rename(Trait = Trait...1) %>%  
  mutate(Trait_Sex = "Male") %>% 
  select(-(Trait...3))

p3 <-
  R_male_traits %>% 
  group_by(Trait) %>% 
  mutate(avg_R = median(Response_to_selection_male)) %>%
  ggplot(aes(Response_to_selection_male, fct_reorder(Trait, avg_R))) +
  stat_interval(.width = c(0.05, 0.66, 0.95), 
                height = 1, show.legend = F) +
  rcartocolor::scale_color_carto_d(palette = "Peach") +
  coord_cartesian(xlim = c(-0.5, 0.5)) +
  geom_vline(linetype = 2, xintercept = 0, linewidth = 1) +
  labs(x = "Response to selection in males",
       y = "Trait expressed in males") +
  theme_bw() + 
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        text = element_text(size=14),
        axis.text.y = element_text(size = 8))

p4 <- 
  R_male_traits %>% 
  group_by(Trait) %>% 
  mutate(avg_R = median(Response_to_selection_female)) %>%
  ggplot(aes(Response_to_selection_female, fct_reorder(Trait, avg_R))) +
  stat_interval(.width = c(0.05, 0.66, 0.95), 
                height = 1, show.legend = F) +
  rcartocolor::scale_color_carto_d(palette = "Purp") +
  coord_cartesian(xlim = c(-0.5, 0.5)) +
  geom_vline(linetype = 2, xintercept = 0, linewidth = 1) +
  labs(x = "Response to selection in females",
       y = "Trait expressed in males") +
  theme_bw() + 
  theme(legend.position = "none",
        panel.grid.minor = element_blank(),
        text = element_text(size=14),
        axis.text.y = element_text(size = 8)) 

p3 + p4 +
  plot_annotation(tag_levels = 'a')

```

**Figure S2**. The response to selection in **a** males and **b** females for all traits measured in males. Innermost bands approximate the median, while outer bands show the 66 and 95% credible intervals.

## Calculate various metrics

$~$

Combine the data frames and estimate the overall expected response to selection as 

$$R = \frac{R_{F} + R_{M}}{2}$$

In the same code chunk, for each trait, we also calculate the difference in the response to selection acting on females and males as $\Delta R = R_{FF} - R_{MF}$ for female traits or $\Delta R = R_{FM} - R_{MM}$ for male traits. This difference facilitates identification of traits where the response to selection is very different between sexes (in magnitude, and possibly also in sign). Note that $\Delta R$ alone does not reveal whether a trait is sexually antagonistic, because sexually concordant selection that varies in strength between sexes (or sex-limited selection) can result in a high value of $\Delta R$. 

```{r}

R_female_traits <-
  R_female_traits %>% 
  mutate(R_overall = (Response_to_selection_female + Response_to_selection_male)/2,
         R_diff = Response_to_selection_female - Response_to_selection_male) %>% 
    select(Trait, Trait_Sex, everything()) 

R_male_traits <-
  R_male_traits %>%  
  mutate(R_overall = (Response_to_selection_female + Response_to_selection_male)/2,
         R_diff = Response_to_selection_female - Response_to_selection_male) %>% 
  select(Trait, Trait_Sex, everything()) %>% 
  filter(Trait != "dopamine.response.to.paraquat.2021.m")

R_all_traits <- bind_rows(R_female_traits, R_male_traits)

R_long_form <-
  R_all_traits %>% 
  select(1:4) %>% 
    pivot_longer(cols = 3:4, names_to = "Fitness_Sex", values_to = "R_metric") %>% 
    mutate(Fitness_Sex = case_when(Fitness_Sex == "Response_to_selection_female" ~ "Female",
                                   Fitness_Sex == "Response_to_selection_male" ~ "Male"))

```

$~$

## Calculating evidence ratios

We use the `brms` `hypothesis()` function to compute evidence ratios, where our hypothesis is $R$ (and its various related metrics) != 0. When the hypothesis is one-sided, this is the posterior probability under the hypothesis against its alternative. That is, when the hypothesis is of the form R > 0, the evidence ratio is the ratio of the posterior probability of R > 0 and the posterior probability of R < 0. In this example, values greater than one indicate that the evidence in favour of R > 0 is larger than evidence in favour of R < 0. In contrast, values smaller than one indicate that there is greater evidence in favour of R < 0 than R > 0. More on the `hypothesis` function can be found [here](https://paul-buerkner.github.io/brms/reference/hypothesis.html). We consider evidence ratios > 39 or < 1/39 as biologically notable (without accounting for multiple testing), as our hypothesis is two-directional (negative and positive values of $R$ are of interest to us) and these values closely correspond to the familiar frequentist p-value = 0.05 [(Makowski et al, 2019)](https://www.frontiersin.org/journals/psychology/articles/10.3389/fpsyg.2019.02767/full).

```{r}

# build a function to calculate the evidence ratio, get the relevant info from the output and convert it to a tibble

find_evidence_ratios <- function(Trait_name, specify_hypothesis){
  x <- hypothesis(R_all_traits %>% 
                    filter(Trait == Trait_name),
                  specify_hypothesis)
  
  x <- x$hypothesis
  
  x %>% as_tibble() %>% 
    select(Evid.Ratio) %>% 
    mutate(Trait = Trait_name)
}

# list of traits we need evidence ratios for.

all_traits_list <- R_all_traits %>% distinct(Trait) %>% pull(Trait)

# calculate evidence ratios 

if(!file.exists("data/organismal_level_output/organismal_level_evidence_ratios.csv")){

R_evidence_ratios_female <- 
  map_dfr(all_traits_list, 
          find_evidence_ratios, 
          "Response_to_selection_female > 0") %>% 
  rename(Female_Evid_Ratio = Evid.Ratio)

R_evidence_ratios_male <-
  map_dfr(all_traits_list, 
          find_evidence_ratios, 
          "Response_to_selection_male > 0") %>% 
  rename(Male_Evid_Ratio = Evid.Ratio)

R_evidence_ratios_overall <- 
    map_dfr(all_traits_list, 
          find_evidence_ratios, 
          "R_overall > 0") %>% 
  rename(Overall_Evid_Ratio = Evid.Ratio)

R_evidence_ratios_diff <- 
    map_dfr(all_traits_list, 
          find_evidence_ratios, 
          "R_diff > 0") %>% 
  rename(Diff_Evid_Ratio = Evid.Ratio)
}
```

Next, we manually calculate evidence ratios for **sexually concordant responses to selection** by:

1. Using the `p_direction` function from the `bayestestR` package, we find the proportion of the posterior distribution (the posterior probability) that is of the median's sign for each trait in each sex.

2. Using the `p_direction` output, we calculate the probability that a trait has positive $R$, $P(pos)$ or negative $R$, $P(neg) = 1 - P(pos)$.

3. Find $P(concord) = P(pos)_f P(pos)_m + P(neg)_f P(neg)_m$

4. Find $P(antag) = P(pos)_f P(neg)_m + P(neg)_f P(pos)_m$

5. Take the ratio of $P(concord)$ and $P(antag)$

Evidence ratios > 1 indicate that the response to selection is more likely to be sexually concordant, whereas ratios < 1 indicate a sexually antagonistic response has higher probability. As indicated by step 3, a concordant response is possible when the trait responds to selection in either a negative or positive direction in both sexes. Similarly, step 4 shows that an antagonistic response is also possible via two paths. Hence, we once again consider evidence ratios > 39 or < 1/39 as biologically notable (without accounting for multiple testing).

```{r}
pd_function <-function(Trait_name){
  R_all_traits %>% filter(Trait == Trait_name) %>%
    select(Trait, Response_to_selection_female, Response_to_selection_male) %>%
    p_direction(Response_to_selection_female, method = "direct", null = 0) %>%
    as_tibble() %>%
    pivot_wider(names_from = Parameter, values_from = pd) %>%
    mutate(Trait = Trait_name) %>%
    rename(pd_female = Response_to_selection_female,
           pd_male = Response_to_selection_male)
}

if(!file.exists("data/organismal_level_output/organismal_level_evidence_ratios.csv")){

pd_data <- map_dfr(all_traits_list, pd_function)

Trait_medians <-
  R_all_traits %>%
  group_by(Trait) %>%
  summarise(median_female = median(Response_to_selection_female),
            median_male = median(Response_to_selection_male)) %>%
  ungroup()

evidence_ratios_concord <-
  left_join(pd_data, Trait_medians) %>%
  mutate(prob_pos_female = if_else(median_female > 0, pd_female, 1 - pd_female),
         prob_pos_male = if_else(median_male > 0, pd_male, 1 - pd_male)) %>%
  select(Trait, prob_pos_female, prob_pos_male) %>%

    # Calculate the probabilities that beta_i and beta_j have the same/opposite signs
    mutate(p_sex_concord = prob_pos_female  * prob_pos_male +
             (1 - prob_pos_female) * (1 - prob_pos_male),
           p_sex_antag = prob_pos_female * (1 - prob_pos_male) +
             (1 - prob_pos_female) * prob_pos_male) %>%
    # Find the ratios of these two probabilities (i.e. the "evidence ratios")
    mutate(Concord_Evid_Ratio = p_sex_concord / p_sex_antag)
}

``` 

Join the evidence ratios into a single tibble and save as a `.csv` file for fast loading. Create new columns that indicate whether the evidence ratio indicate biologically notability.

```{r}
if(!file.exists("data/organismal_level_output/organismal_level_evidence_ratios.csv")){
organismal_level_evidence_ratios <-
  R_evidence_ratios_female %>% 
  left_join(R_evidence_ratios_male) %>% 
  left_join(R_evidence_ratios_overall) %>% 
  left_join(R_evidence_ratios_diff) %>% 
  left_join(evidence_ratios_concord %>% select(Trait, Concord_Evid_Ratio)) %>% 
  mutate(Female_notable = if_else(Female_Evid_Ratio > 39 | Female_Evid_Ratio < 1/39, "Yes", "No"),
         Male_notable = if_else(Male_Evid_Ratio > 39 | Male_Evid_Ratio < 1/39, "Yes", "No"),
         Overall_notable = if_else(Overall_Evid_Ratio > 39 | Overall_Evid_Ratio < 1/39, "Yes", "No"),
         Diff_notable = if_else(Diff_Evid_Ratio > 39 | Diff_Evid_Ratio < 1/39, "Yes", "No"),
         Concord_notable = if_else(Concord_Evid_Ratio > 39 | Concord_Evid_Ratio < 1/39, "Yes", "No")) %>% 
  left_join(clean_meta_data) %>% 
  left_join(R_all_traits %>% distinct(Trait, Trait_Sex)) %>%
  mutate(Trait = fct_reorder(Trait, `Trait guild`)) %>%
  arrange(`Trait guild`, Trait) %>% 
  mutate(position = 1:n(),
         Trait_Sex = if_else(Trait_Sex == "Female", "Traits measured in females",
                                         "Traits measured in males")) %>% 
  select(Trait, everything())

write_csv(evidence_ratios, "data/organismal_level_output/organismal_level_evidence_ratios.csv")
} else organismal_level_evidence_ratios <- read_csv("data/organismal_level_output/organismal_level_evidence_ratios.csv")

```

$~$

Build the evidence ratio plots

```{r, fig.height=11, fig.width=10}

# work out where to place the x axis labels - we want them in the middle of their trait guild

x_labels <- organismal_level_evidence_ratios %>% 
  group_by(`Trait guild`) %>% 
  summarise(position = min(position) + (max(position) - min(position))/2)

# get colours for each trait guild

guild_pal <- c("#6e7cb9", "#7bbcd5", "#d0e2af", "#f5db99",
               "#e89c81", "#d2848d", "#6e7cb9", "#7bbcd5", 
               "#d0e2af","#f5db99","#e89c81", "#d2848d", 
               "#6e7cb9", "#7bbcd5")

plot_a <- 
  organismal_level_evidence_ratios %>% 
  ggplot(aes(x = position, y = log2(Overall_Evid_Ratio))) +
  geom_point(aes(colour = `Trait guild`),
             size = 4.5, alpha = 1, shape = 17) +
  geom_hline(yintercept = log2(1), linetype = 2, linewidth = 1) +
  geom_hline(yintercept = log2(39), linetype = 3, linewidth = 0.9) +
  geom_hline(yintercept = log2(1/39), linetype = 3, linewidth = 0.9) +
  #scale_shape_manual(values = c(25, 24)) +
  scale_colour_manual(values = guild_pal) +
  scale_x_continuous(breaks = x_labels$position, labels = x_labels$`Trait guild`) +
  scale_y_continuous(expand = c(0.01, 0), breaks = c(-12, -10, -8, -6, -4, -2, 0, 2, 4, 6, 8, 10, 12), 
                  labels = c(paste("1/",2 ^ c(12, 10, 8, 6, 4, 2), sep = ""), 2 ^ c(0, 2,4,6,8,10,12)), limits = c(-14, 14)) +
  labs(x = NULL,
       #y = "R~F~ log2(ER)",
       y = "_R_  log~2~(ER)") +
  geom_label_repel(data = organismal_level_evidence_ratios %>% 
                     filter(Overall_Evid_Ratio > 200 | Overall_Evid_Ratio < 1/200),  
                   aes(label = Trait_nice),
                   fill = "white",
                   size = 3, alpha = 0.9,
                   min.segment.length = 0, seed = 1,
                   box.padding = 0.5, point.padding = 0.5,
                   max.overlaps = 30) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_markdown(size = 15))
    
plot_b <- 
  organismal_level_evidence_ratios %>% 
  ggplot(aes(x = position, y = log2(Female_Evid_Ratio))) +
  geom_point(aes(colour = `Trait guild`),
             size = 4.5, alpha = 1, shape = 17) +
  geom_hline(yintercept = log2(1), linetype = 2, linewidth = 1) +
  geom_hline(yintercept = log2(39), linetype = 3, linewidth = 0.9) +
  geom_hline(yintercept = log2(1/39), linetype = 3, linewidth = 0.9) +
  #scale_shape_manual(values = c(25, 24)) +
  scale_colour_manual(values = guild_pal) +
  scale_x_continuous(breaks = x_labels$position, labels = x_labels$`Trait guild`) +
  scale_y_continuous(expand = c(0.01, 0), breaks = c(-12, -10, -8, -6, -4, -2, 0, 2, 4, 6, 8, 10, 12), 
                  labels = c(paste("1/",2 ^ c(12, 10, 8, 6, 4, 2), sep = ""), 2 ^ c(0, 2,4,6,8,10,12)), limits = c(-14, 14)) +
  labs(x = NULL,
       #y = "R~F~ log2(ER)",
       y = "_R_ on females log~2~(ER)") +
  geom_label_repel(data = organismal_level_evidence_ratios %>% 
                     filter(Female_Evid_Ratio > 100 | Female_Evid_Ratio < 1/100),  
                   aes(label = Trait_nice),
                   fill = "white",
                   size = 3, alpha = 0.9,
                   min.segment.length = 0, seed = 1,
                   box.padding = 0.5, point.padding = 0.5,
                   max.overlaps = 30) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_markdown(size = 15))

plot_c <- 
  organismal_level_evidence_ratios %>% 
  ggplot(aes(x = position, y = log2(Male_Evid_Ratio))) +
  geom_point(aes(colour = `Trait guild`),
             size = 4.5, alpha = 1, shape = 17) +
  geom_hline(yintercept = log2(1), linetype = 2, linewidth = 1) +
  geom_hline(yintercept = log2(39), linetype = 3, linewidth = 0.9) +
  geom_hline(yintercept = log2(1/39), linetype = 3, linewidth = 0.9) +
  #scale_shape_manual(values = c(25, 24)) +
  scale_colour_manual(values = guild_pal) +
  scale_x_continuous(breaks = x_labels$position, labels = x_labels$`Trait guild`) +
  scale_y_continuous(expand = c(0.01, 0), breaks = c(-12, -10, -8, -6, -4, -2, 0, 2, 4, 6, 8, 10, 12), 
                  labels = c(paste("1/",2 ^ c(12, 10, 8, 6, 4, 2), sep = ""), 2 ^ c(0, 2,4,6,8,10,12)), limits = c(-14, 14)) +
  labs(x = "Traits",
       y = "_R_ on males log~2~(ER)") +
  geom_label_repel(data = organismal_level_evidence_ratios %>% 
                     filter(Male_Evid_Ratio > 100 | Male_Evid_Ratio < 1/100),
                   aes(label = Trait_nice),
                   fill = "white",
                   size = 3, alpha = 0.9,
                   min.segment.length = 0, seed = 1,
                   box.padding = 0.5, point.padding = 0.5,
                   max.overlaps = 30) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x=element_text(angle = 90, 
                                 vjust = 0.5,
                                 size = 10),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_markdown(size = 15))

(Figure_S3 <- plot_a / plot_b / plot_c +
  plot_annotation(tag_levels = 'a'))

```

**Figure S3**. Triangles show evidence ratios (ERs) for **a** overall $R$, **b** $R$ on females ($R_{FF}$ and $R_{MF}$) and **c** $R$ on males ($R_{FM}$ and $R_{MM}$). Traits are loosely organised into guilds to aid visualisation. The dashed lines indicate an evidence ratio of 1, where the probability that $R$ > 0 is equal to the probability that $R$ < 0. ER > 1 indicates a positive response is more likely, whereas ER < 1 indicates a negative response is more likely. The upper dotted line on each plot indicates an evidence ratio of 39, while the lower dotted line indicates an evidence ratio of 1/39; these correspond strongly with the frequentist $p$ = 0.05. Traits with the most extreme evidence ratios are labelled.

```{r, fig.height=9, fig.width=12.5}
plot_d <- 
  organismal_level_evidence_ratios %>% 
  ggplot(aes(x = position, y = log2(Diff_Evid_Ratio))) +
  geom_point(aes(colour = `Trait guild`),
             size = 4.5, alpha = 1, shape = 17) +
  geom_hline(yintercept = log2(1), linetype = 2, linewidth = 1) +
  geom_hline(yintercept = log2(39), linetype = 3, linewidth = 0.9) +
  geom_hline(yintercept = log2(1/39), linetype = 3, linewidth = 0.9) +
  scale_colour_manual(values = guild_pal) +
  scale_x_continuous(breaks = x_labels$position, labels = x_labels$`Trait guild`) +
  scale_y_continuous(expand = c(0.01, 0), breaks = c(-10, -8, -6, -4, -2, 0, 2, 4, 6, 8, 10), 
                     labels = c(paste("1/",2 ^ c(10, 8, 6, 4, 2), sep = ""), 2 ^ c(0, 2,4,6,8,10)), limits = c(-12, 12)) +
  labs(x = NULL,
       y = "&Delta;_R_  log~2~(ER)") +
  geom_label_repel(data = organismal_level_evidence_ratios %>% 
                     mutate(Trait_nice = case_when(Trait_nice == "DDT resistance mortality" ~ "DDT susceptibility", .default = Trait_nice)) %>% 
                     filter(Diff_Evid_Ratio > 39 | Diff_Evid_Ratio < 1/39),  
                   aes(label = Trait_nice),
                   fill = "white",
                   size = 3, alpha = 0.9,
                   min.segment.length = 0, seed = 1,
                   box.padding = 0.5, point.padding = 0.5,
                   max.overlaps = 30) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_markdown(size = 15))

plot_e <- 
  organismal_level_evidence_ratios %>% 
  ggplot(aes(x = position, y = log2(Concord_Evid_Ratio))) +
  geom_point(aes(colour = `Trait guild`),
             size = 4.5, alpha = 1, shape = 17) +
  geom_hline(yintercept = log2(1), linetype = 2, linewidth = 1) +
  geom_hline(yintercept = log2(39), linetype = 3, linewidth = 0.9) +
  geom_hline(yintercept = log2(1/39), linetype = 3, linewidth = 0.9) +
  scale_colour_manual(values = guild_pal) +
  scale_x_continuous(breaks = x_labels$position, labels = x_labels$`Trait guild`) +
  scale_y_continuous(expand = c(0.01, 0), breaks = c(-10, -8, -6, -4, -2, 0, 2, 4, 6, 8, 10), 
                  labels = c(paste("1/",2 ^ c(10, 8, 6, 4, 2), sep = ""), 2 ^ c(0, 2,4,6,8,10)), limits = c(-12, 12)) +
  labs(x = "Trait",
       y = "Sexual concordance log~2~(ER)") +
  geom_label_repel(data = organismal_level_evidence_ratios %>% 
                     filter(Concord_Evid_Ratio > 39 | Concord_Evid_Ratio < 1/39),  
                   aes(label = Trait_nice),
                   fill = "white",
                   size = 3, alpha = 0.9,
                   min.segment.length = 0, seed = 1,
                   box.padding = 0.5, point.padding = 0.5,
                   max.overlaps = 30) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x=element_text(angle = 90, 
                                 vjust = 0.5,
                                 size = 10),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_markdown(size = 15))

(Figure_3 <- plot_d/ plot_e + plot_annotation(tag_levels = "a"))
```

**Figure 3**. Triangles show evidence ratios (ERs) for **a** $\Delta R$ ($R$ on females - $R$ on males) and **b** sexual concordance, for all measured traits in the organismal level phenotype dataset. Traits are loosely organised into guilds to aid visualisation. The dashed lines indicate an evidence ratio of 1, where the probability that $\Delta R$ > 0 is equal to the probability that $\Delta R$ < 0, or that a trait is just as likely to have a sexually concordant response to selection (ER > 1) as a sexually antagonistic response (ER < 1). The upper dotted line on each plot indicates an evidence ratio of 39, while the lower dotted line indicates an evidence ratio of 1/39; these correspond strongly with the frequentist $p$ = 0.05. Traits with the most extreme evidence ratios are labelled.

## Table S2

**Table S2** / **Supplementary dataset S1**: $R$ partitions and associated evidence ratios for the organismal level trait dataset. Trait descriptions are provided in Table S1. Data can be downloaded as a `.csv` file from the [github repository](https://github.com/tomkeaney/DGRP_sexual_conflict) associated with this project.
```{r}
all_ol_traits_summary <-
  R_all_traits %>% 
  group_by(Trait, Trait_Sex) %>% 
  median_qi(Response_to_selection_female,
            Response_to_selection_male,
            R_overall,
            R_diff) %>% 
  rename(`Sex trait was measured in` = Trait_Sex,
         R = R_overall,
         `R 2.5% CI` = R_overall.lower,
         `R 97.5% CI` = R_overall.upper,
         `R female` = Response_to_selection_female,
         `R female 2.5% CI` = Response_to_selection_female.lower,
         `R female 97.5% CI` = Response_to_selection_female.upper,
         `R male` = Response_to_selection_male,
         `R male 2.5% CI` = Response_to_selection_male.lower,
         `R male 97.5% CI` = Response_to_selection_male.upper,
         `Delta R` = R_diff,
         `Delta R 2.5% CI` = R_diff.lower,
         `Delta R 97.5% CI` = R_diff.upper) %>% 
  left_join(organismal_level_evidence_ratios %>% select(-contains("notable")) %>% 
              rename(`R ER` = Overall_Evid_Ratio, `R female ER` = Female_Evid_Ratio,
                     `R male ER` = Male_Evid_Ratio, `Delta R ER` = Diff_Evid_Ratio,
                     `Sexual concordance ER` = Concord_Evid_Ratio)) %>% 
  select(Trait, Trait_nice, `Sex trait was measured in`, R, `R 2.5% CI`, `R 97.5% CI`, `R ER`, 
         `R female`, `R female 2.5% CI`, `R female 97.5% CI`, `R female ER`,
         `R male`, `R male 2.5% CI`, `R male 97.5% CI`, `R male ER`,
         `Delta R`, `Delta R 2.5% CI`, `Delta R 97.5% CI`, `Delta R ER`, `Sexual concordance ER`) %>%
  ungroup() %>% 
  mutate(across(c(4,5,6,8,9,10,12,13,14,16,17,18), ~round(.x, 3)),
         across(ends_with("ER"), ~ round(.x, 4)))

write_csv(all_ol_traits_summary, "data/organismal_level_output/Supplementary_dataset_1.csv")

my_data_table(all_ol_traits_summary)
  
```
              
$~$

## Estimate $\overline{|R|}$ in each sex

### The model

Fit the model to test whether $|R|$ is affected by the sex fitness and trait values were measured in. Here, we take a meta-analysis type approach and weight each $R$ estimate by the inverse of its standard error. 


```{r}

R_medians <-
  R_long_form %>%
  group_by(Trait, Fitness_Sex, Trait_Sex) %>% 
  summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>% 
  ungroup() %>% 
  select(-variable) %>% 
  left_join(clean_meta_data) %>% 
  mutate(absolute_R = abs(median),
         Fitness_Sex = as.factor(Fitness_Sex),
         Trait_Sex = as.factor(Trait_Sex),
         Trait = as.factor(Trait)) 

median_R_model <- 
  brm(absolute_R | weights(1/sd) ~ 1 + Fitness_Sex * Trait_Sex + (1|study_ID) + (1|Trait),
      family = brmsfamily(family = "Gamma"), # gamma is appropriate for the half-normal distribution created by taking the absolute
      data = R_medians,
      prior = c(prior(normal(-2.2, 1), class = Intercept),
                prior(exponential(1), class = sd),
                prior(exponential(1), class = shape),
                prior(normal(0, 0.25), class = b)),
      warmup = 4000, iter = 12000,
      seed = 1, cores = 4, chains = 4,
      control = list(adapt_delta = 0.9, max_treedepth = 10),
      file = "fits/median_R_model")

print(median_R_model)

```

To check how does the gamma distribution fares at capturing the shape of the data, we use the model to recapitulate the data it was trained on. Given the shape of the predicted data is similar to the real data, there's a high chance we've used an appropriate distribution family.

```{r}
pp_check(median_R_model)
```

$~$

### Build Panels a and b for Figure 2

Get model predictions and plot

```{r, fig.width=10, fig.height=5}

new_data <- expand_grid(Fitness_Sex = c("Female", "Male"),
                        Trait_Sex = c("Female", "Male"))

R_fitted <-
  fitted(median_R_model, newdata = new_data, re_formula = NA, summary = F) %>% 
  as.data.frame() %>% 
  rename(FemaleFitness_FemaleTrait = V1, FemaleFitness_MaleTrait = V2, 
         MaleFitness_FemaleTrait = V3, MaleFitness_MaleTrait = V4) %>% 
  as_tibble() %>% 
  mutate(percent_diff_female = ((MaleFitness_FemaleTrait - FemaleFitness_FemaleTrait) / FemaleFitness_FemaleTrait)*100,
         percent_diff_male = ((MaleFitness_MaleTrait - FemaleFitness_MaleTrait)/ FemaleFitness_MaleTrait)*100) %>% 
  pivot_longer(cols = everything(), names_to = "Parameter", values_to = "R_mean") %>% 
  mutate(Fitness_Sex = case_when(str_detect(Parameter, "FemaleFitness") ~ "Female",
                                 str_detect(Parameter, "MaleFitness") ~ "Male"),
         Trait_Sex = case_when(str_detect(Parameter, "FemaleTrait") ~ "Female",
                               str_detect(Parameter, "MaleTrait") ~ "Male",
                               str_detect(Parameter, "percent_diff_female") ~ "Female",
                               str_detect(Parameter, "percent_diff_male") ~ "Male"))

R_p1 <-
  R_fitted %>% 
  filter(!str_detect(Parameter, "percent")) %>% 
  ggplot(aes(x = R_mean, y = Trait_Sex, fill = Fitness_Sex)) + #fill = Fitness_Sex)) +
  stat_slab(alpha = 0.9, shape = 21) +#,
  labs(x = expression("Mean |"* italic(R) * "| for organismal level traits"),
       y = "Phenotyped sex") +
  scale_fill_manual(values = c(carto_pal(7, "Purp")[5], carto_pal(7, "Peach")[5])) +
  coord_cartesian(xlim = c(0.045, 0.10)) +
  theme_minimal() + 
  theme(panel.background = element_rect(fill='transparent'),
        #panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.background = element_rect(fill='transparent', color=NA),
        legend.position = "none",
        text = element_text(size=13))

R_p2 <-
  R_fitted %>% 
  filter(str_detect(Parameter, "percent")) %>%
  ggplot(aes(x = R_mean, y = Trait_Sex, fill = after_stat(x > 0))) +
  stat_halfeye(.width = c(0.66, 0.95), alpha = 0.9, fill = carto_pal(7, "Emrld")[2],
               point_interval = "median_qi", point_fill = "white",
               shape = 21, point_size = 4, stroke = 1.2) + 
  geom_vline(xintercept = 0, linetype = 2, linewidth = 1.2) +
  #coord_cartesian(xlim = c(-5, 50)) +
  xlab(expression("% sex difference in |"*italic(R)* "|")) +
  ylab("Phenotyped sex") +
  coord_cartesian(xlim = c(-5, 45)) +
  theme_minimal() + 
  theme(panel.background = element_rect(fill='transparent'),
        #panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.background = element_rect(fill='transparent', color=NA),
        legend.position = "none",
        text = element_text(size=13))

R_p1 + R_p2 +
  plot_annotation(tag_levels = 'a')
  
```

$~$

**Calculating stats for the results section**

```{r}
R_fitted %>% 
  group_by(Parameter) %>% 
  median_qi(R_mean)

```

$~$

# Transcriptome analysis

## Selecting data appropriate for analysis

We use the data archived on the dgrp2 [website](http://dgrp2.gnets.ncsu.edu/data.html), which was provided by [Huang et al. (2015)](https://pubmed.ncbi.nlm.nih.gov/26483487/). These data are levels of expression of genes measured across 185 DGRP lines. Huang et al's data contains Y-linked genes that have higher/equal expression in *females* in all lines, presumably microarray issues/errors. To be conservative, we restrict our analyses to genes that are known to be on chromosomes that are present in both sexes. After data cleaning, we retain 14,268 genes in our analysis.

We also load gene annotation data using the `org.Dm.eg.db` R package provided by `BiocManager`. The code used to produce annotations is provided in the `code` subdirectory, in the `get_gene_annotations.R` file. 

Huang et al. 2015 replicated each gene expression measurement twice. To find line means, we simply take the average value for each line. We then standardise the expression of each gene to have $\mu = 0$ and $\sigma = 1$ (as with the 'phenotypic' traits.

```{r}

# load in the gene names, we'll use these for plots and tables

Dmel_gene_names <- read_csv("data/input/all_dmel_genes.csv")


gene_annotations <- read_csv("data/input/gene_anntotations.csv")


# Helper to load all the Huang et al. expression data into tidy format
load_expression_data <- function(gene_annotations){
  
  # Note: Huang et al's data contains Y-linked genes that have
  # higher/equal expression in *females* in all lines, presumably microarray issues/errors.
  # To be conservative, we restrict our analyses to genes that are known to be on a
  # chromosomes that is present in both sexes
  genes_allowed <- gene_annotations %>%
    filter(chromosome %in% c("2L", "2R", "3L", "3R", "4", "X")) %>%
    pull(FBID) 
  
  females <- read_delim("data/input/huang_transcriptome/dgrp.array.exp.female.txt", delim = " ") %>%
    filter(gene %in% genes_allowed) %>% 
    gather(sample, expression, -gene) %>% 
    mutate(line = map_chr(str_extract_all(sample, "line_[:digit:]*"), ~ .x[1]),
           replicate = map_chr(str_split(sample, ":"), ~ .x[2]),
           sex = "Female")
  
  males <- read_delim("data/input/huang_transcriptome/dgrp.array.exp.male.txt", delim = " ") %>%
    filter(gene %in% genes_allowed) %>% 
    gather(sample, expression, -gene) %>% 
    mutate(line = map_chr(str_extract_all(sample, "line_[:digit:]*"), ~ .x[1]),
           replicate = map_chr(str_split(sample, ":"), ~ .x[2]),
           sex = "Male")
  
  bind_rows(females, males) %>% 
    select(sample, line, sex, replicate, gene, expression)
}

expression_line_means <- 
  load_expression_data(gene_annotations) %>% # use the custom function to load expression data
  mutate(line = str_remove(line, "line_"),
         line = as.factor(line)) %>% 
  group_by(line, sex, gene) %>% 
  mutate(trait_value = mean(expression)) %>% # compute the average between replicates for each gene
  ungroup() %>% 
  distinct(line, sex, gene, trait_value) %>% 
  group_by(sex, gene) %>% # scale the traits, specific to gene and sex
  mutate(trait_value = as.numeric(scale(trait_value))) %>% 
  ungroup() %>% 
  rename(Sex = sex) %>% 
# join the fitness data
left_join(trait_data %>% distinct(line, female_fitness, male_fitness))

# the transcriptome is large; memory is therefore a consistent problem with this analysis - it helps to clear often
# 
invisible(gc())

```

$~$

## Calculating _R_: the response to selection

$~$

### Build models to calculate $R_{FF}$ & $R_{MF}$

To estimate the covariance between $A_w$ and $A_z$ (which in this case is a transcript), we fitted bivariate Bayesian linear models using the `brms` package (Bürkner, 2017) for `R version 4.2.2`. For each combination of trait/transcript and sex, we used line means for the focal trait/transcript and the fitness of the focal sex as the two response variables and fitted an intercept-only Gaussian model. Each model returned a posterior distribution of the residual correlation between trait/transcript and fitness, which for data expressed in standard units is equivalent to the covariance.

Build functions to run the models

```{r}

# RFF estimates

female_transcripts <- expression_line_means %>% filter(Sex == "Female")

transcript_list <- unique(female_transcripts$gene) # an input to the map_dfr() function that we'll need in a few chunks time

# code the model structure we will use for all traits/transcripts using one example - `FBgn0000014`. We can then use the update() function to run this model many times, once for each trait/transcript measured in females. update() makes this process many times faster, because the model can immediately start sampling, without the need to recompile. 

RFF_transcriptome_model <-
  brm(data = female_transcripts %>% filter(gene == "FBgn0000014"),
      family = gaussian,
      bf(mvbind(female_fitness, trait_value) ~ 1) + set_rescor(TRUE),
      prior = c(prior(normal(0, 0.1), class = Intercept, resp = femalefitness),
                prior(normal(0, 0.1), class = Intercept, resp = traitvalue),
                prior(normal(1, 0.1), class = sigma, resp = femalefitness),
                prior(normal(1, 0.1), class = sigma, resp = traitvalue),
                prior(lkj(2), class = rescor)),
      chains = 4, cores = 4, iter = 4000, warmup = 2000,
      seed = 1, file = "fits/RFF_transcriptome_model",
      backend = "cmdstanr", refresh = 400)  


# make a function to update the model and the posterior sample output with the 'selected trait'

RFF_transcriptome_calculator <- function(selected_gene){
  
  data <- female_transcripts %>% filter(gene == selected_gene)
  
  model <- update(
    RFF_transcriptome_model, newdata = data,
      chains = 4, cores = 4, iter = 4000, warmup = 2000,
      seed = 1, backend = "cmdstanr", refresh = 400)
  
  posterior <- 
    as_draws_df(model) %>% 
    rename(Response_to_selection_female = rescor__femalefitness__traitvalue) %>% 
    mutate(Trait = selected_gene) %>% 
    select(Trait, Response_to_selection_female) %>% 
    as_tibble()
  
  posterior
}

# RMF estimates

RMF_transcriptome_model <-
  brm(data = female_transcripts %>% filter(gene == "FBgn0000014"),
      family = gaussian,
      bf(mvbind(male_fitness, trait_value) ~ 1) + set_rescor(TRUE),
      prior = c(prior(normal(0, 0.1), class = Intercept, resp = malefitness),
                prior(normal(0, 0.1), class = Intercept, resp = traitvalue),
                prior(normal(1, 0.1), class = sigma, resp = malefitness),
                prior(normal(1, 0.1), class = sigma, resp = traitvalue),
                prior(lkj(2), class = rescor)),
      chains = 4, cores = 4, iter = 4000, warmup = 2000,
      seed = 1, file = "fits/RMF_transcriptome_model",
      backend = "cmdstanr", refresh = 400) 

# make a function to update the model and the posterior sample output with your desired trait

RMF_transcriptome_calculator <- function(selected_gene){
  
  data <- female_transcripts %>% filter(gene == selected_gene)
  
  model <- update(
    RMF_transcriptome_model, newdata = data,
      chains = 4, cores = 4, iter = 4000, warmup = 2000,
      seed = 1, backend = "cmdstanr", refresh = 400)
  
  posterior <- 
    as_draws_df(model) %>% 
    rename(Response_to_selection_male = rescor__malefitness__traitvalue) %>% 
    mutate(Trait = selected_gene) %>% 
    select(Trait, Response_to_selection_male) %>% 
    as_tibble()
  
  posterior
}

```

$~$

### Build models to calculate $R_{FM}$ & $R_{MF}$

$~$

```{r}

# RMM estimates

male_transcripts <- expression_line_means %>% filter(Sex == "Male")

RMM_transcriptome_model <-
  brm(data = male_transcripts %>% filter(gene == "FBgn0000014"),
      family = gaussian,
      bf(mvbind(male_fitness, trait_value) ~ 1) + set_rescor(TRUE),
      prior = c(prior(normal(0, 0.1), class = Intercept, resp = malefitness),
                prior(normal(0, 0.1), class = Intercept, resp = traitvalue),
                prior(normal(1, 0.1), class = sigma, resp = malefitness),
                prior(normal(1, 0.1), class = sigma, resp = traitvalue),
                prior(lkj(2), class = rescor)),
      chains = 4, cores = 4, iter = 4000, warmup = 2000,
      seed = 1, file = "fits/RMM_transcriptome_model",
      backend = "cmdstanr", refresh = 400)    

# make a function to update the model and the posterior sample output with your desired trait

RMM_transcriptome_calculator <- function(selected_gene){
  
  data <- male_transcripts %>% filter(gene == selected_gene)
  
  model <- update(
    RMM_transcriptome_model, newdata = data,
    chains = 4, cores = 4, iter = 4000, warmup = 2000,
    seed = 1)
  
  posterior <- 
    as_draws_df(model) %>% 
    rename(Response_to_selection_male = rescor__malefitness__traitvalue) %>%
    mutate(Trait = selected_gene) %>% 
    select(Trait, Response_to_selection_male) %>% 
    as_tibble()
  
  posterior
}

# RFM estimates

RFM_transcriptome_model <-
  brm(data = male_transcripts %>% filter(gene == "FBgn0000014"),
      family = gaussian,
      bf(mvbind(female_fitness, trait_value) ~ 1) + set_rescor(TRUE),
      prior = c(prior(normal(0, 0.1), class = Intercept, resp = femalefitness),
                prior(normal(0, 0.1), class = Intercept, resp = traitvalue),
                prior(normal(1, 0.1), class = sigma, resp = femalefitness),
                prior(normal(1, 0.1), class = sigma, resp = traitvalue),
                prior(lkj(2), class = rescor)),
      chains = 4, cores = 4, iter = 4000, warmup = 2000,
      seed = 1, file = "fits/RFM_transcriptome_model",
      backend = "cmdstanr", refresh = 400)    

# make a function to update the model and the posterior sample output with your desired trait

RFM_transcriptome_calculator <- function(selected_trait){
  
  data <- male_transcripts %>% filter(gene == selected_trait)
  
  model <- update(
    RFM_transcriptome_model, newdata = data,
      chains = 4, cores = 4, iter = 4000, warmup = 2000,
      seed = 1)
  
  posterior <- 
    as_draws_df(model) %>% 
    rename(Response_to_selection_female = rescor__femalefitness__traitvalue) %>% 
    mutate(Trait = selected_gene) %>% 
    select(Trait, Response_to_selection_female) %>% 
    as_tibble()
  
  posterior
}
```

$~$

### Run the models for all the traits

Run the models using `RFF_transcriptome_calculator`, `RMF_transcriptome_calculator`, `RMM_transcriptome_calculator` and `RFM_transcriptome_calculator`.

This takes a fair bit of memory, so it might be necessary to run the models in chunks rather than all in one go. To do this, you can break the `transcript_list_female` list into parts and feed it into the `map_dfr` function. The completed subset can then be saved to your hard disk, removed from R and cleared from your computers memory. This frees up memory to run another chunk without losing progress. Expand the code chunk below to see an example. All other chunks have been run and saved for later use.

```{r}
# run the RFF function

transcript_list_1 <- transcript_list[1:2000]

if(!file.exists("data/transcriptome_output/transcriptome_chunks/RFF_transcript_1.csv")){
  RFF_transcript_1 <- map_dfr(transcript_list_1, RFF_transcriptome_calculator)
  write_csv(RFF_transcript_1, 
            file = "data/transcriptome_output/transcriptome_chunks/RFF_transcript_1.csv")
  rm(RFF_transcript_1)
  invisible(gc())
} else RFF_transcript_1 <- read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_1.csv")

```

Load all the posterior draws, combine and summarise and save the result to the hard disk. This allows us to simply load the summarised results once everything has been run once.

```{r}

if(!file.exists("data/transcriptome_output/R_summarised_transcriptome.csv")){
RFF_transcriptome_complete <- 
  rbind(
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_1.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_2.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_3.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_4.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_5.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_6.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_7.csv")
  ) %>% 
  group_by(Trait) %>% 
  summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
  ungroup() %>% 
  select(-variable) %>% 
  mutate(absolute_R = abs(median),
         Fitness_Sex = "Female",
         Trait_Sex = "Female")

invisible(gc())

RMF_transcriptome_complete <- 
  rbind(
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_1.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_2.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_3.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_4.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_5.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_6.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_7.csv")
  ) %>% 
  group_by(Trait) %>% 
  summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
  ungroup() %>% 
  select(-variable) %>% 
  mutate(absolute_R = abs(median),
         Fitness_Sex = "Male",
         Trait_Sex = "Female")

invisible(gc())

RMM_transcriptome_complete <- 
  rbind(
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_1.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_2.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_3.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_4.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_5.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_6.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_7.csv")
  ) %>% 
  group_by(Trait) %>% 
  summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
  ungroup() %>% 
  select(-variable) %>% 
  mutate(absolute_R = abs(median),
         Fitness_Sex = "Male",
         Trait_Sex = "Male")

invisible(gc())

RFM_transcriptome_complete <- 
  rbind(
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_1.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_2.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_3.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_4.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_5.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_6.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_7.csv")
  ) %>% 
  group_by(Trait) %>% 
  summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
  ungroup() %>% 
  select(-variable) %>% 
  mutate(absolute_R = abs(median),
         Fitness_Sex = "Female",
         Trait_Sex = "Male")

invisible(gc())

R_summarised_transcriptome <-
  bind_rows(RFF_transcriptome_complete, RMF_transcriptome_complete,
            RFM_transcriptome_complete, RMM_transcriptome_complete)

write_csv(R_summarised_transcriptome, "data/transcriptome_output/R_summarised_transcriptome.csv")

} else R_summarised_transcriptome <- read_csv("data/transcriptome_output/R_summarised_transcriptome.csv")

```

$~$

## Calculate various metrics

First, find $R_{overall}$

```{r}
if(!file.exists("data/transcriptome_output/R_overall_transcript_summarised.csv")){
  R_overall_transcript_summarised <-
    bind_rows(
      left_join(read_csv(
        "data/transcriptome_output/transcriptome_chunks/RFF_transcript_1.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv(
                  "data/transcriptome_output/transcriptome_chunks/RMF_transcript_1.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),
      
      left_join(read_csv(
        "data/transcriptome_output/transcriptome_chunks/RFF_transcript_2.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv(
                  "data/transcriptome_output/transcriptome_chunks/RMF_transcript_2.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>%
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),
      
      left_join(read_csv(
        "data/transcriptome_output/transcriptome_chunks/RFF_transcript_3.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv(
                  "data/transcriptome_output/transcriptome_chunks/RMF_transcript_3.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>%
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),
      
      left_join(read_csv(
        "data/transcriptome_output/transcriptome_chunks/RFF_transcript_4.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv(
                  "data/transcriptome_output/transcriptome_chunks/RMF_transcript_4.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>%
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),

      left_join(read_csv(
        "data/transcriptome_output/transcriptome_chunks/RFF_transcript_5.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv(
                  "data/transcriptome_output/transcriptome_chunks/RMF_transcript_5.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>%
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),

      left_join(read_csv(
        "data/transcriptome_output/transcriptome_chunks/RFF_transcript_6.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv(
                  "data/transcriptome_output/transcriptome_chunks/RMF_transcript_6.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),

      left_join(read_csv(
        "data/transcriptome_output/transcriptome_chunks/RFF_transcript_7.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv(
                  "data/transcriptome_output/transcriptome_chunks/RMF_transcript_7.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>%
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),
      
      # now the traits measured in males
      
      left_join(read_csv(
        "data/transcriptome_output/transcriptome_chunks/RMM_transcript_1.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_1.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),

      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_2.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_2.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),

      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_3.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_3.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>%
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),

      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_4.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_4.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>%
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_5.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_5.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>%
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_6.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_6.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>%
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),

      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_7.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_7.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_overall =(Response_to_selection_female + Response_to_selection_male)/2) %>%
        select(Trait, R_overall) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_overall = median) %>%
        select(-variable) %>% 
        mutate(Trait_Sex = "Male")
    )
  
  write_csv(R_overall_transcript_summarised, "data/transcriptome_output/R_overall_transcript_summarised.csv")
} else R_overall_transcript_summarised <- read_csv("data/transcriptome_output/R_overall_transcript_summarised.csv")

```

Now find $\Delta R$

```{r}
if(!file.exists("data/transcriptome_output/R_diff_transcript_summarised.csv")){
  R_diff_transcript_summarised <-
    bind_rows(
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_1.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_1.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_2.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_2.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_3.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_3.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
                mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),

      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_4.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_4.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),

      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_5.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_5.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),
 
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_6.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_6.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>%  
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),

      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_7.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_7.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Female"),
      
      # now the traits measured in males
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_1.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_1.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_2.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_2.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_3.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_3.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_4.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_4.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_5.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_5.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_6.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_6.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Male"),
      
      left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_7.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(),
                read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_7.csv") %>% 
                  group_by(Trait) %>% 
                  mutate(draw = row_number()) %>% 
                  ungroup(), by = c("Trait", "draw")) %>% 
        mutate(R_diff = Response_to_selection_female - Response_to_selection_male) %>%
        select(Trait, R_diff) %>% 
        group_by(Trait) %>% 
        summarise_draws("median", "sd", ~quantile(.x, probs = c(0.025, 0.975), na.rm = TRUE)) %>%
        ungroup() %>% 
        rename(R_diff = median) %>% 
        select(-variable) %>% 
        mutate(Trait_Sex = "Male")
    )
  
  write_csv(R_diff_transcript_summarised, "data/transcriptome_output/R_diff_transcript_summarised.csv")
} else R_diff_transcript_summarised <- read_csv("data/transcriptome_output/R_diff_transcript_summarised.csv")

```

## Calculating evidence ratios

Use the `brms` `hypothesis()` function to compute evidence ratios, just as we did for the organismal level phenotypic traits. 

```{r}
# build a function to calculate the evidence ratio, get the relevant info from the output and convert it to a tibble. Basically the same function, just with a different dataframe input and a memory clearing step that slows it down but allows it to run for many traits. 

find_evidence_ratios_transcriptome <- function(Trait_name, specify_hypothesis){
  x <- hypothesis(R_transcriptome %>% 
                    filter(Trait == Trait_name),
                  specify_hypothesis)
  
  x <- x$hypothesis
  
  invisible(gc())
  
  x %>% as_tibble() %>% 
    select(Evid.Ratio) %>% 
    mutate(Trait = Trait_name)
}

```

This pc has 16gb memory; not enough to load in all the transcript posterior draws at once. Having previously saved these to the hard disk we load the draws for all female measured traits and calculate the evidence ratio that $R_{FF} > 0$, $R_{MF} > 0$, $R_F > 0$ and $\Delta R > 0$. The draws for the female measured traits were then removed from the computers memory using the `rm()` and `gc()` functions and the process was repeated for the male measured traits.

```{r}

# find evidence ratios for the traits measured in females

if(!file.exists("data/transcriptome_output/evidence_ratios_transcriptome_female.csv")){ # name refers to traits measured in females

R_transcriptome <-
  cbind(
  rbind(
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_1.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_2.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_3.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_4.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_5.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_6.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_7.csv")
  ),
   rbind(
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_1.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_2.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_3.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_4.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_5.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_6.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_7.csv")) %>% 
      select(-Trait)
   )

R_transcriptome <- 
  R_transcriptome %>% 
  mutate(R_overall = (Response_to_selection_female - Response_to_selection_male) / 2,
         R_diff = Response_to_selection_female - Response_to_selection_male) 
  
# calculate evidence ratios 

R_evidence_ratios_female <- 
  map_dfr(transcript_list, 
          find_evidence_ratios_transcriptome, 
          "Response_to_selection_female > 0") %>% 
  rename(Female_Evid_Ratio = Evid.Ratio)

R_evidence_ratios_male <-
  map_dfr(transcript_list, 
          find_evidence_ratios_transcriptome, 
          "Response_to_selection_male > 0") %>% 
  rename(Male_Evid_Ratio = Evid.Ratio)

R_evidence_ratios_overall <- 
    map_dfr(transcript_list, 
          find_evidence_ratios_transcriptome, 
          "R_overall > 0") %>% 
  rename(Overall_Evid_Ratio = Evid.Ratio)

R_evidence_ratios_diff <- 
    map_dfr(transcript_list, 
          find_evidence_ratios_transcriptome, 
          "R_diff > 0") %>% 
  rename(Diff_Evid_Ratio = Evid.Ratio)

# write each result to csv in 

evidence_ratios_transcriptome_female <-
  R_evidence_ratios_female %>% 
  left_join(R_evidence_ratios_male) %>% 
  left_join(R_evidence_ratios_overall) %>% 
  left_join(R_evidence_ratios_diff) %>% 
  select(Trait, everything()) %>% 
  rename(Female_Evid_Ratio = Female_evidence_ratio) %>%  # fixing small typo
  mutate(Trait_Sex = "Female")

write_csv(evidence_ratios_transcriptome_female, "data/transcriptome_output/evidence_ratios_transcriptome_female.csv")

} else evidence_ratios_transcriptome_female <- read_csv("data/transcriptome_output/evidence_ratios_transcriptome_female.csv")

```

```{r}

# find evidence ratios for the traits measured in males

if(!file.exists("data/transcriptome_output/evidence_ratios_transcriptome_male.csv")){ # name refers to traits measured in females

R_transcriptome <-
  cbind(
  rbind(
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_1.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_2.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_3.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_4.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_5.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_6.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_7.csv")
  ),
   rbind(
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_1.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_2.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_3.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_4.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_5.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_6.csv"),
    read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_7.csv")) %>% 
      select(-Trait)
   )

R_transcriptome <- 
  R_transcriptome %>% 
  mutate(R_overall = (Response_to_selection_female - Response_to_selection_male) / 2,
         R_diff = Response_to_selection_female - Response_to_selection_male) 
  
# calculate evidence ratios 

R_evidence_ratios_female <- 
  map_dfr(transcript_list, 
          find_evidence_ratios_transcriptome, 
          "Response_to_selection_female > 0") %>% 
  rename(Female_Evid_Ratio = Evid.Ratio)

R_evidence_ratios_male <-
  map_dfr(transcript_list, 
          find_evidence_ratios_transcriptome, 
          "Response_to_selection_male > 0") %>% 
  rename(Male_Evid_Ratio = Evid.Ratio)

R_evidence_ratios_overall <- 
    map_dfr(transcript_list, 
          find_evidence_ratios_transcriptome, 
          "R_overall > 0") %>% 
  rename(Overall_Evid_Ratio = Evid.Ratio)

R_evidence_ratios_diff <- 
    map_dfr(transcript_list, 
          find_evidence_ratios_transcriptome, 
          "R_diff > 0") %>% 
  rename(Diff_Evid_Ratio = Evid.Ratio)

evidence_ratios_transcriptome_male <-
  R_evidence_ratios_female %>% 
  left_join(R_evidence_ratios_male) %>% 
  left_join(R_evidence_ratios_overall) %>% 
  left_join(R_evidence_ratios_diff) %>% 
  select(Trait, everything()) %>% 
  rename(Female_Evid_Ratio = Female_evidence_ratio) %>%  # fixing small typo
  mutate(Trait_Sex = "Male")

write_csv(evidence_ratios_transcriptome_male, "data/transcriptome_output/evidence_ratios_transcriptome_male.csv")

} else evidence_ratios_transcriptome_male <- read_csv("data/transcriptome_output/evidence_ratios_transcriptome_male.csv")

```

To calculate evidence ratios for sexually concordant responses to selection we use the same function as for the organismal-level phenotypic traits. Run the function in chunks, save them to hard-disk, clear memory and repeat. If already done, just load.

```{r}

if(!file.exists("data/transcriptome_output/concord_evidence_ratio_data.csv")){

female_transcripts_1 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_1.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_1.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

trait_list_1 <- unique(female_transcripts_1$Trait)

f1 <- get_evidence_ratios(female_transcripts_1, trait_list_1, "Female")

rm(female_transcripts_1)
invisible(gc())

female_transcripts_2 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_2.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_2.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

trait_list_2 <- unique(female_transcripts_2$Trait)

f2 <- get_evidence_ratios(female_transcripts_2, trait_list_2, "Female")

rm(female_transcripts_2)
invisible(gc())

female_transcripts_3 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_3.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_3.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

trait_list_3 <- unique(female_transcripts_3$Trait)

f3 <- get_evidence_ratios(female_transcripts_3, trait_list_3, "Female")

rm(female_transcripts_3)
invisible(gc())

female_transcripts_4 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_4.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_4.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

trait_list_4 <- unique(female_transcripts_4$Trait)

f4 <- get_evidence_ratios(female_transcripts_4, trait_list_4, "Female")

rm(female_transcripts_4)
invisible(gc())

female_transcripts_5 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_5.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_5.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

trait_list_5 <- unique(female_transcripts_5$Trait)

f5 <- get_evidence_ratios(female_transcripts_5, trait_list_5, "Female")

rm(female_transcripts_5)
invisible(gc())

female_transcripts_6 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_6.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_6.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

trait_list_6 <- unique(female_transcripts_6$Trait)

f6 <- get_evidence_ratios(female_transcripts_6, trait_list_6, "Female")

rm(female_transcripts_6)
invisible(gc())

female_transcripts_7 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RFF_transcript_7.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RMF_transcript_7.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

trait_list_7 <- unique(female_transcripts_7$Trait)

f7 <- get_evidence_ratios(female_transcripts_7, trait_list_7, "Female")

rm(female_transcripts_7)
invisible(gc())

# now the transcripts measured in males

male_transcripts_1 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_1.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_1.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

m_trait_list_1 <- unique(male_transcripts_1$Trait)

m1 <- get_evidence_ratios(male_transcripts_1, m_trait_list_1, "Male")

rm(male_transcripts_1)
invisible(gc())

male_transcripts_2 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_2.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_2.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

m_trait_list_2 <- unique(male_transcripts_2$Trait)

m2 <- get_evidence_ratios(male_transcripts_2, m_trait_list_2, "Male")

rm(male_transcripts_2)
invisible(gc())

male_transcripts_3 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_3.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_3.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

m_trait_list_3 <- unique(male_transcripts_3$Trait)

m3 <- get_evidence_ratios(male_transcripts_3, m_trait_list_3, "Male")

rm(male_transcripts_3)
invisible(gc())

male_transcripts_4 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_4.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_4.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

m_trait_list_4 <- unique(male_transcripts_4$Trait)

m4 <- get_evidence_ratios(male_transcripts_4, m_trait_list_4, "Male")

rm(male_transcripts_4)
invisible(gc())

male_transcripts_5 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_5.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_5.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

m_trait_list_5 <- unique(male_transcripts_5$Trait)

m5 <- get_evidence_ratios(male_transcripts_5, m_trait_list_5, "Male")

rm(male_transcripts_5)
invisible(gc())

male_transcripts_6 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_6.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_6.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

m_trait_list_6 <- unique(male_transcripts_6$Trait)

m6 <- get_evidence_ratios(male_transcripts_6, m_trait_list_6, "Male")

rm(male_transcripts_6)
invisible(gc())

male_transcripts_7 <-
  left_join(read_csv("data/transcriptome_output/transcriptome_chunks/RMM_transcript_7.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(),
            read_csv("data/transcriptome_output/transcriptome_chunks/RFM_transcript_7.csv") %>% 
              group_by(Trait) %>% 
              mutate(draw = row_number()) %>% 
              ungroup(), by = c("Trait", "draw"))

m_trait_list_7 <- unique(male_transcripts_7$Trait)

m7 <- get_evidence_ratios(male_transcripts_7, m_trait_list_7, "Male")

rm(male_transcripts_7)
invisible(gc())

concord_evidence_ratio_data <- 
  bind_rows(f1, f2, f3, f4, f5, f6, f7, m1, m2, m3, m4, m5, m6, m7)

write_csv(concord_evidence_ratio_data, "data/transcriptome_output/concord_evidence_ratio_data.csv")

} else concord_evidence_ratio_data <- read_csv("data/transcriptome_output/concord_evidence_ratio_data.csv")

```

Bind all the evidence ratios together

```{r}

evidence_ratios_transcriptome <-
  evidence_ratios_transcriptome_female %>% 
  bind_rows(evidence_ratios_transcriptome_male %>% rename(Male_Evid_Ratio = Male_evidence_ratio)) %>%  # fix a typo
  left_join(concord_evidence_ratio_data %>% rename(Concord_Evid_Ratio = evidence_ratio_concord) %>% select(Trait, Trait_Sex, Concord_Evid_Ratio),
            by = c("Trait", "Trait_Sex")) %>% 
  left_join(gene_annotations %>% rename(Trait = FBID)) %>% 
  mutate(chromosome_numeric = case_when(chromosome == "2L" ~ 1,
                                        chromosome == "2R" ~ 2,
                                        chromosome == "3L" ~ 3,
                                        chromosome == "3R" ~ 4,
                                        chromosome == "X" ~ 5)) %>% 
  #left_join(R_summarised_transcriptome %>% distinct(Trait, Trait_Sex)) %>%
  #mutate(Trait = fct_reorder(Trait, chromosome_numeric))
  arrange(chromosome, Trait) %>% 
  mutate(position = 1:n(),
         Trait_Sex = if_else(Trait_Sex == "Female", "Traits measured in females",
                                         "Traits measured in males")) %>% 
  select(Trait, everything())

```

$~$

Find the number of evidence ratios > 39 or < 1/39 for each partition of $R$

```{r}
response_pos <- evidence_ratios_transcriptome %>% filter(Overall_Evid_Ratio > 39) %>% arrange(-Overall_Evid_Ratio)

response_neg <- evidence_ratios_transcriptome %>% filter(Overall_Evid_Ratio < 1/39) %>% arrange(Overall_Evid_Ratio)

response_to_female_pos <- evidence_ratios_transcriptome %>% filter(Female_Evid_Ratio > 39) %>% arrange(-Female_Evid_Ratio)

response_to_female_neg <- evidence_ratios_transcriptome %>% filter(Female_Evid_Ratio < 1/39) %>% arrange(Female_Evid_Ratio)

response_to_male_pos <- evidence_ratios_transcriptome %>% filter(Male_Evid_Ratio > 39) %>% arrange(-Male_Evid_Ratio)

response_to_male_neg <- evidence_ratios_transcriptome %>% filter(Male_Evid_Ratio < 1/39) %>% arrange(Male_Evid_Ratio)

response_sexes_diff <- 
  evidence_ratios_transcriptome %>% filter(Diff_Evid_Ratio > 39 | Diff_Evid_Ratio < 1/39) %>% arrange(-Diff_Evid_Ratio)

sexually_concordant <- evidence_ratios_transcriptome %>% filter(Concord_Evid_Ratio > 39) %>% arrange(-Concord_Evid_Ratio)

sexually_antagonistic <- evidence_ratios_transcriptome %>% filter(Concord_Evid_Ratio < 1/39) %>% arrange(Concord_Evid_Ratio)
```

`r nrow(response_pos)` transcripts are predicted to have a positive overall response to selection, while `r nrow(response_neg)` transcripts ar expected to have a negative overall response. 

In response to selection on females, `r nrow(response_to_female_pos)` transcripts are expected to increase in expression, whereas `r nrow(response_to_female_neg)` are expected to decrease in mean expression.

In response to selection on males, `r nrow(response_to_male_pos)` transcripts are expected to increase in expression, whereas `r nrow(response_to_male_neg)` are expected to decrease in mean expression.

`r nrow(response_sexes_diff)` transcripts have notably different responses to selection on females compared with males.  

`r nrow(sexually_antagonistic)` transcripts are expected to have a sexually antagonistic response to selection, while `r nrow(sexually_concordant)` transcripts are expected to have a sexually concordant response. From these numbers, we can roughly estimate that 3.8% of transcripts that are responding to selection in both sexes are sexually antagonistic. 

Build the evidence ratio plots

```{r, fig.height=8, fig.width=11}

x_labels <- evidence_ratios_transcriptome %>% 
  group_by(chromosome) %>% 
  summarise(position = min(position) + (max(position) - min(position))/2)


guild_pal <- c("#6e7cb9", "#7bbcd5", "#d0e2af", "#f5db99",
               "#e89c81", "#d2848d")

plot_transcriptome_a <- 
  evidence_ratios_transcriptome %>% 
  ggplot(aes(x = position, y = log2(Overall_Evid_Ratio))) +
  geom_point(aes(colour = chromosome),
             size = 2.5, alpha = 0.8, shape = 17) +
  geom_hline(yintercept = log2(1), linetype = 2, linewidth = 1) +
  geom_hline(yintercept = log2(39), linetype = 3, linewidth = 0.9) +
  geom_hline(yintercept = log2(1/39), linetype = 3, linewidth = 0.9) +
  #scale_shape_manual(values = c(25, 24)) +
  scale_colour_manual(values = guild_pal) +
  scale_x_continuous(breaks = x_labels$position, labels = x_labels$chromosome) +
  scale_y_continuous(expand = c(0.01, 0), breaks = c(-14, -12, -10, -8, -6, -4, -2, 0, 2, 4, 6, 8, 10, 12, 14), 
                     labels = c(paste("1/",2 ^ c(14, 12, 10, 8, 6, 4, 2), sep = ""), 2 ^ c(0, 2,4,6,8,10,12,14)), limits = c(-18, 18)) +
  labs(x = NULL,
       #y = "R~F~ log2(ER)",
       y = "_R_  log~2~(ER)") +
  geom_label_repel(data = evidence_ratios_transcriptome %>% 
                     filter(Overall_Evid_Ratio > 1000 | Overall_Evid_Ratio < 1/1000 ),  
                   aes(label = gene_symbol),
                   fill = "white",
                   size = 3, alpha = 0.9,
                   min.segment.length = 0, seed = 1,
                   box.padding = 0.5, point.padding = 0.5,
                   max.overlaps = 30) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_markdown(size = 10))

plot_transcriptome_b <- 
  evidence_ratios_transcriptome %>% 
  ggplot(aes(x = position, y = log2(Female_Evid_Ratio))) +
  geom_point(aes(colour = chromosome),
             size = 2.5, alpha = 0.8, shape = 17) +
  geom_hline(yintercept = log2(1), linetype = 2, linewidth = 1) +
  geom_hline(yintercept = log2(39), linetype = 3, linewidth = 0.9) +
  geom_hline(yintercept = log2(1/39), linetype = 3, linewidth = 0.9) +
  #scale_shape_manual(values = c(25, 24)) +
  scale_colour_manual(values = guild_pal) +
  scale_x_continuous(breaks = x_labels$position, labels = x_labels$chromosome) +
  scale_y_continuous(expand = c(0.01, 0), breaks = c(-14, -12, -10, -8, -6, -4, -2, 0, 2, 4, 6, 8, 10, 12, 14), 
                     labels = c(paste("1/",2 ^ c(14, 12, 10, 8, 6, 4, 2), sep = ""), 2 ^ c(0, 2,4,6,8,10,12,14)), limits = c(-16, 16)) +
  labs(x = NULL,
       #y = "R~F~ log2(ER)",
       y = "_R_ on females log~2~(ER)") +
  geom_label_repel(data = evidence_ratios_transcriptome %>% 
                     filter(Female_Evid_Ratio > 1000 | Female_Evid_Ratio < 1/1000 ), 
                   aes(label = gene_symbol),
                   fill = "white",
                   size = 3, alpha = 0.9,
                   min.segment.length = 0, seed = 1,
                   box.padding = 0.5, point.padding = 0.5,
                   max.overlaps = 30) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_markdown(size = 10))

plot_transcriptome_c <- 
  evidence_ratios_transcriptome %>% 
  ggplot(aes(x = position, y = log2(Male_Evid_Ratio))) +
   geom_point(aes(colour = chromosome),
             size = 2.5, alpha = 0.8, shape = 17) +
  geom_hline(yintercept = log2(1), linetype = 2, linewidth = 1) +
  geom_hline(yintercept = log2(39), linetype = 3, linewidth = 0.9) +
  geom_hline(yintercept = log2(1/39), linetype = 3, linewidth = 0.9) +
  #scale_shape_manual(values = c(25, 24)) +
  scale_colour_manual(values = guild_pal) +
  scale_x_continuous(breaks = x_labels$position, labels = x_labels$chromosome) +
  scale_y_continuous(expand = c(0.01, 0), breaks = c(-14, -12, -10, -8, -6, -4, -2, 0, 2, 4, 6, 8, 10, 12, 14), 
                     labels = c(paste("1/",2 ^ c(14, 12, 10, 8, 6, 4, 2), sep = ""), 2 ^ c(0, 2,4,6,8,10,12,14)), limits = c(-15, 15)) +
  labs(x = "Chromosome arm",
       #y = "R~F~ log2(ER)",
       y = "_R_ on males log~2~(ER)") +
  geom_label_repel(data = evidence_ratios_transcriptome %>% 
                     filter(Male_Evid_Ratio > 1000 | Male_Evid_Ratio < 1/1000 ),  
                   aes(label = gene_symbol),
                   fill = "white",
                   size = 3, alpha = 0.9,
                   min.segment.length = 0, seed = 1,
                   box.padding = 0.5, point.padding = 0.5,
                   max.overlaps = 30) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x=element_text(size = 15),
        #axis.ticks.x=element_blank(),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_markdown(size = 10))

(Figure_S4 <- plot_transcriptome_a / plot_transcriptome_b / plot_transcriptome_c +
  plot_annotation(tag_levels = 'a')
)

```

**Figure S4**. Triangles show evidence ratios (ERs) for gene expression traits. **a** shows overall $R$, **b** $R$ on females ($R_{FF}$ and $R_{MF}$) and **c** $R$ on males ($R_{FM}$ and $R_{MM}$). The dashed lines indicate an evidence ratio of 1, where the probability that $R$ > 0 is equal to the probability that $R$ < 0. ER > 1 indicates a positive response is more likely, whereas ER < 1 indicates a negative response is more likely. The upper dotted line on each plot indicates an evidence ratio of 39, while the lower dotted line indicates an evidence ratio of 1/39; these correspond strongly with the frequentist $p$ = 0.05. Traits with the most extreme evidence ratios are labelled.

```{r, fig.height=7, fig.width=11}
plot_transcriptome_d <- 
  evidence_ratios_transcriptome %>% 
  ggplot(aes(x = position, y = log2(Diff_Evid_Ratio))) +
   geom_point(aes(colour = chromosome),
             size = 2.5, alpha = 0.8, shape = 17) +
  geom_hline(yintercept = log2(1), linetype = 2, linewidth = 1) +
  geom_hline(yintercept = log2(39), linetype = 3, linewidth = 0.9) +
  geom_hline(yintercept = log2(1/39), linetype = 3, linewidth = 0.9) +
  #scale_shape_manual(values = c(25, 24)) +
  scale_colour_manual(values = guild_pal) +
  scale_x_continuous(breaks = x_labels$position, labels = x_labels$chromosome) +
  scale_y_continuous(expand = c(0.01, 0), breaks = c(-12, -10, -8, -6, -4, -2, 0, 2, 4, 6, 8, 10, 12), 
                     labels = c(paste("1/",2 ^ c(12, 10, 8, 6, 4, 2), sep = ""), 2 ^ c(0, 2,4,6,8,10,12)), limits = c(-13, 13)) +
  labs(x = "Chromosome arm",
       #y = "R~F~ log2(ER)",
       y = "&Delta;_R_  log~2~(ER)") +
  geom_label_repel(data = evidence_ratios_transcriptome %>% 
                     filter(Diff_Evid_Ratio > 250 | Diff_Evid_Ratio < 1/250 ),  
                   aes(label = gene_symbol),
                   fill = "white",
                   size = 3, alpha = 0.9,
                   min.segment.length = 0, seed = 1,
                   box.padding = 0.5, point.padding = 0.5,
                   max.overlaps = 30) +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.x = element_blank(),
        axis.title.y = element_markdown(size = 15))

plot_transcriptome_e <- 
  evidence_ratios_transcriptome %>% 
  ggplot(aes(x = position, y = log2(Concord_Evid_Ratio))) +
   geom_point(aes(colour = chromosome),
             size = 2.5, alpha = 0.8, shape = 17) +
  geom_hline(yintercept = log2(1), linetype = 2, linewidth = 1) +
  geom_hline(yintercept = log2(39), linetype = 3, linewidth = 0.9) +
  geom_hline(yintercept = log2(1/39), linetype = 3, linewidth = 0.9) +
  #scale_shape_manual(values = c(25, 24)) +
  scale_colour_manual(values = guild_pal) +
  scale_x_continuous(breaks = x_labels$position, labels = x_labels$chromosome) +
  scale_y_continuous(expand = c(0.01, 0), breaks = c(-12, -10, -8, -6, -4, -2, 0, 2, 4, 6, 8, 10, 12), 
                     labels = c(paste("1/",2 ^ c(12, 10, 8, 6, 4, 2), sep = ""), 2 ^ c(0, 2,4,6,8,10,12)), limits = c(-12, 12)) +
  labs(x = "Chromosome arm",
       y = "Sexual concrdoance  log~2~(ER)") +
  geom_label_repel(data = evidence_ratios_transcriptome %>% 
                     filter(Concord_Evid_Ratio > 39 | Concord_Evid_Ratio < 1/39), 
                   aes(label = gene_symbol),
                   fill = "white",
                   size = 3, alpha = 0.9,
                   min.segment.length = 0, seed = 1,
                   box.padding = 0.5, point.padding = 0.5,
                   max.overlaps = 30) +
  theme_minimal() +
    theme(legend.position = "none",
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.x=element_text(size = 15),
        #axis.ticks.x=element_blank(),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_markdown(size = 15))


(Figure_4 <- plot_transcriptome_d/ plot_transcriptome_e + plot_annotation(tag_levels = "a")
)
```

**Figure 4**. Triangles show evidence ratios (ERs) for **a** $\Delta R$ ($R$ on females - $R$ on males) and **b** sexual concordance, for all measured traits in the transcriptome dataset. The dashed lines indicate an evidence ratio of 1, where the probability that $\Delta R$ > 0 is equal to the probability that $\Delta R$ < 0, or that a trait is just as likely to have a sexually concordant response to selection (ER > 1) as a sexually antagonistic response (ER < 1). The upper dotted line on each plot indicates an evidence ratio of 39, while the lower dotted line indicates an evidence ratio of 1/39; these correspond strongly with the frequentist $p$ = 0.05. Traits with the most extreme evidence ratios are labelled.

$~$

## Table S3

**Table S3** / **Supplementary dataset S2**: $R$ partitions and associated evidence ratios for the gene expression trait dataset. Data can be downloaded as a `.csv` file from the [github repository](https://github.com/tomkeaney/DGRP_sexual_conflict) associated with this project.
```{r}

all_transcriptome_summary <-
  R_summarised_transcriptome %>% 
  select(-c(sd, absolute_R)) %>% 
  pivot_wider(names_from = c(Fitness_Sex), values_from = 2:4) %>% 
  rename(`R female` = median_Female,
         `R female 2.5% CI` = `2.5%_Female`,
         `R female 97.5% CI` = `97.5%_Female`,
         `R male` = median_Male,
         `R male 2.5% CI` = `2.5%_Male`,
         `R male 97.5% CI` = `97.5%_Male`) %>% 
  left_join(
    
    R_overall_transcript_summarised %>%
      select(-sd) %>% 
      rename( R = R_overall,
              `R 2.5% CI` = `2.5%`,
              `R 97.5% CI` = `97.5%`),
    by = c("Trait", "Trait_Sex")) %>% 
  left_join(
    R_diff_transcript_summarised %>%
      select(-sd) %>% 
      rename(`Delta R` = R_diff,
             `Delta R 2.5% CI` = `2.5%`,
             `Delta R 97.5% CI` = `97.5%`),
    by = c("Trait", "Trait_Sex")) %>% 
  
  left_join(evidence_ratios_transcriptome %>% 
              mutate(Trait_Sex = case_when(Trait_Sex == "Traits measured in females" ~ "Female",
                                           Trait_Sex == "Traits measured in males" ~ "Male")) %>% 
              select(-c(entrez_id, chromosome, chromosome_numeric, position)) %>% 
              rename(`R ER` = Overall_Evid_Ratio, `R female ER` = Female_Evid_Ratio,
                     `R male ER` = Male_Evid_Ratio, `Delta R ER` = Diff_Evid_Ratio,
                     `Sexual concordance ER` = Concord_Evid_Ratio)) %>% 
  rename(`Sex trait was measured in` = Trait_Sex) %>% 
    select(Trait, gene_name, gene_symbol, `Sex trait was measured in`, R, `R 2.5% CI`, `R 97.5% CI`, `R ER`, 
           `R female`, `R female 2.5% CI`, `R female 97.5% CI`, `R female ER`,
           `R male`, `R male 2.5% CI`, `R male 97.5% CI`, `R male ER`,
           `Delta R`, `Delta R 2.5% CI`, `Delta R 97.5% CI`, `Delta R ER`, `Sexual concordance ER`) %>% 
  mutate(across(c(5,6,7,9,10,11,13,14,15,17,18,19), ~round(.x, 3)),
         across(ends_with("ER"), ~ round(.x, 4))) 

write_csv(all_transcriptome_summary, "data/transcriptome_output/Supplementary_dataset_2.csv")

my_data_table(all_transcriptome_summary)
  
```
              
$~$

## Estimate $\overline{|R|}$ in each sex

$~$

### The model

Fit the model to test whether $\overline{|R|}$ depends on the sex fitness and trait values were measured in:

```{r}

# fit the model

median_R_transcriptome_model <- 
  brm(absolute_R | weights(1/sd) ~ 1 + Fitness_Sex * Trait_Sex + (1|Trait),
      family = brmsfamily(family = "Gamma"), # gamma is appropriate for the half-normal distribution created by taking the absolute
      data = R_summarised_transcriptome,
      prior = c(prior(normal(-2.2, 1), class = Intercept),
                prior(exponential(1), class = sd),
                prior(exponential(1), class = shape),
                prior(normal(0, 0.25), class = b)),
      warmup = 2000, iter = 6000,
      seed = 1, cores = 4, chains = 4,
      control = list(adapt_delta = 0.8, max_treedepth = 10),
      file = "fits/median_R_transcriptome_model")

print(median_R_transcriptome_model)

invisible(gc())
```

To check how does the gamma distribution fares at capturing the shape of the data, we use the model to recapitulate the data it was trained on. Given the shape of the predicted data is similar to the real data, there's a high chance we've used an appropriate distribution family.

```{r}
pp_check(median_R_transcriptome_model)
```

$~$

Get model predictions. Once again, we only do this once and save for later convenience.

```{r, fig.width=10, fig.height=5}

new_data <- expand_grid(Fitness_Sex = c("Female", "Male"),
                        Trait_Sex = c("Female", "Male"))

if(!file.exists("data/transcriptome_output/R_transcriptome_fitted.csv")){
R_transcriptome_fitted <-
  fitted(median_R_transcriptome_model, newdata = new_data, re_formula = NA, summary = F) %>% 
  as.data.frame() %>% 
  rename(FemaleFitness_FemaleTrait = V1, FemaleFitness_MaleTrait = V2, 
         MaleFitness_FemaleTrait = V3, MaleFitness_MaleTrait = V4) %>% 
  as_tibble() %>% 
  mutate(percent_diff_female = 
           ((MaleFitness_FemaleTrait - FemaleFitness_FemaleTrait) / FemaleFitness_FemaleTrait)*100,
         percent_diff_male = 
           ((MaleFitness_MaleTrait - FemaleFitness_MaleTrait)/ FemaleFitness_MaleTrait)*100) %>% 
  pivot_longer(cols = everything(), names_to = "Parameter", values_to = "R_mean") %>% 
  mutate(Fitness_Sex = case_when(str_detect(Parameter, "FemaleFitness") ~ "Female",
                                 str_detect(Parameter, "MaleFitness") ~ "Male"),
         Trait_Sex = case_when(str_detect(Parameter, "FemaleTrait") ~ "Female",
                               str_detect(Parameter, "MaleTrait") ~ "Male",
                               str_detect(Parameter, "percent_diff_female") ~ "Female",
                               str_detect(Parameter, "percent_diff_male") ~ "Male"))

write_csv(R_transcriptome_fitted, "data/transcriptome_output/R_transcriptome_fitted.csv")
} else R_transcriptome_fitted <- read_csv("data/transcriptome_output/R_transcriptome_fitted.csv")
  
```

$~$

### Complete Figure 2

Create panels b and d

```{r}
R_t1 <-
  R_transcriptome_fitted %>% 
  filter(!str_detect(Parameter, "percent")) %>% 
  ggplot(aes(x = R_mean, y = Trait_Sex, fill = Fitness_Sex)) + #+ y = Trait_Sex)) + #fill = Fitness_Sex)) +
  stat_slab(alpha = 0.9, shape = 21) +#,
  labs(x = expression("Mean |"* italic(R) * "| for gene expression traits"),
       y = "Phenotyped sex",
       fill = "Sex under selection") +
  scale_fill_manual(values = c(carto_pal(7, "Purp")[5], carto_pal(7, "Peach")[5])) +
  coord_cartesian(xlim = c(0.045, 0.10)) +
  theme_minimal() + 
  theme(panel.background = element_rect(fill='transparent'),
        #panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.background = element_rect(fill='transparent', color=NA),
        legend.position = "bottom",
        text = element_text(size=13))

R_t2 <-
  R_transcriptome_fitted %>% 
  filter(str_detect(Parameter, "percent")) %>%
  ggplot(aes(x = R_mean, y = Trait_Sex)) + #, #y = Trait_Sex, fill = Trait_Sex)) +
  stat_halfeye(.width = c(0.66, 0.95), alpha = 0.9, fill = carto_pal(7, "Emrld")[2],
               point_interval = "median_qi", point_fill = "white",
               shape = 21, point_size = 4, stroke = 1.2) + # width indicates the uncertainty intervals: here we have 66% and 95% intervals + # width indicates the uncertainty intervals: here we have 66% and 95% intervals+
  #scale_fill_manual(values = c(carto_pal(7, "Peach")[5], carto_pal(7, "Peach")[5], carto_pal(7, "Purp")[5])) +
  geom_vline(xintercept = 0, linetype = 2, linewidth = 1.2) +
  coord_cartesian(xlim = c(-5, 45)) +
  xlab(expression("% sex difference in |"*italic(R)* "|")) +
  ylab("Phenotyped sex") +
  theme_minimal() + 
  theme(panel.background = element_rect(fill='transparent'),
        #panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.background = element_rect(fill='transparent', color=NA),
        legend.position = "none",
        text = element_text(size=13))

```

Produce the complete figure

```{r}

left_col <- (R_p1 / R_t1)

right_col <- (R_p2 / R_t2)

wrap_plots(left_col, right_col) & 
  plot_annotation(tag_levels = "a") 
```

**Figure 2**. **a** the posterior distribution for |$\overline{R}$| across organismal level traits, the mean expected response to selection (absolute value) in females and males and split by the sex the trait was measured in. **b** shows |$\overline{R}$| across gene-expression traits. |$\overline{R}$| is expressed in trait standard deviations. **c** and **d** show the percentage increase or decrease in |$\overline{R}$| through males compared to through females for the organismal level and gene expression traits, respectively. Points indicate the estimated mean with associated 66 and 95% credible intervals (where wide enough to be visible).

**Calculating stats for the results section**

```{r}
R_transcriptome_fitted %>% 
  group_by(Parameter) %>% 
  median_qi(R_mean)
```
