---
title: "'SNP-heritabilities' and the Robertson-Price Identity"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE, warning = FALSE, results='hide'
)
```

## Load packages and prepare for `GCTA` analysis

Here we load the necessary packages to build this document and code in 'tidy' style. We also set the file path so that R can access the `gcta64` software. Finally, we build a helper function to pass `GCTA` commands to the terminal. This function allows us to code entirely within R-studio, which is our strong preference.

```{r}
library(tidyverse)
library(glue)
library(kableExtra)
library(DT)
library(MetBrewer)
library(patchwork)
library(rcartocolor)

gcta <- file.path(getwd(), "code/gcta64")
options(readr.show_col_types = FALSE)


# helper function to pass commands to the terminal
# Note that we set `intern = TRUE`, and pass the result of `system()` to `cat()`,
# ensuring that the Terminal output will be printed in this knitr report.
run_command <- function(shell_command, wd = getwd(), path = ""){
  cat(system(glue("cd ", wd, path, "\n",shell_command), intern = TRUE), sep = '\n')
}


### useful code that I didn't end up needing

  #remove the suffix that denotes the sex the trait was measured in and replace the level 'both' with 'pooled'
 # mutate(Sex = recode(Sex, Both = "Pooled"),
  #       Trait = case_when(
   #        str_ends(Trait, ".f") | str_ends(Trait, ".m")  ~ gsub('.{2}$', '', Trait)),
    #     `Trait 2` = case_when(
     #      str_ends(`Trait 2`, ".f") | str_ends(`Trait 2`, ".m")  ~ gsub('.{2}$', '', `Trait 2`))) %>% 

```

## Load phenotype data

```{r}
traits <- read_csv("data/all.dgrp.phenos_unscaled.csv") 

traits_for_gwas <- 
  traits %>% 
  # filter(!(`Trait guild` %in% allowed_guilds) | is.na(`Trait guild`)) %>% 
  group_by(Trait) %>% 
  summarise(lines_measured = length(unique(line))) %>% 
  ungroup() %>% 
  filter(lines_measured > 79) %>%
  filter(Trait != "paraquat.resistance.2021.f" & Trait != "starvation.resistance.high.carb.diet.m") %>% # these traits call a gcta error, we remove for now
  pull(Trait)  # find traits with 80+ replicates
 

traits_for_gwas <- traits_for_gwas[!(grepl("fitness.late", traits_for_gwas))]

# replace any slashes in the trait name, e.g. "1/3-Methylhistidine.high.yeast.f", as this is not ok to use as a file name
traits_for_gwas <- str_replace_all(traits_for_gwas, "[/]", "_") 

line_mean_phenotypes <- read_csv("data/all.dgrp.phenos_unscaled.csv") %>%
  mutate(Trait = str_replace_all(Trait, "[/]", "_")) %>%  # change to match `traits_for_gwas` list
  filter(Trait %in% traits_for_gwas) %>%
   filter(Trait != "paraquat.resistance.2021.f" & Trait != "starvation.resistance.high.carb.diet.m") %>% # these traits produce a gcta error, we remove for now
  select(line, Trait, trait_value) %>% 
  #distinct(line, Trait, .keep_all = TRUE) %>% # Remove this line once Tom resolves the duplicates
  spread(Trait, trait_value) %>% 
  mutate_at(vars(-line), ~ as.numeric(scale(.x))) %>%  # scale the traits
  mutate(line = paste("line", line, sep = "")) # add "line" in front of the line number to fit with plink/gcta formatting
```


## Use `GCTA` to estimate the genetic variances and heritabilities for each trait

Here we estimate 'SNP-heritability' for each phenotype using a genomic relatedness matrix Restricted Maximum Likelihood (GREML) approach. GREML uses a set of SNPs $S$ that are known across a population of individuals (or lines, in our case) to estimate the relatedness between random individuals (lines) and compares that with phenotypic variation across that set of individuals. This allows us to estimate the variation in phenotype that is explained by the known set of SNPs in our population. We can then divide this variance by the total variation found for the phenotype, giving the SNP-heritability. Note that SNP-heritability should be smaller than other estimates of heritability such as broad- or narrow-sense heritability, as it is limited to additive effects from a subset of the genome, while the latter two estimates encompass a broader range of genetic effects. 

We use the `GCTA` software to run this analysis. 

A few helpful exerts from the literature about GREML are provided below:

> "Data from a GWAS that are collected to detect statistical associations between SNPs and complex traits are usually analyzed by testing each SNP individually for an association with the trait. To account for the large number of significance tests carried out, a very stringent P value is used. This reduces the occurrence of false positives, but it may cause many real associations to be missed, especially if individual SNPs have a small effect on the trait." 

`r tufte::quote_footer('--- Yang et al 2010')`

> "In contrast to single-SNP association analysis, the basic concept behind the `gcta64` method is to fit the effects of all the SNPs as random effects using a mixed linear model (MLM)."

`r tufte::quote_footer('--- Yang et al 2011')`

### Make a genomic relatedness matrix for all the lines

This uses the `GCTA` command `--make-grm` to make a genomic relatedness matrix (GRM) for the DGRP SNP data. This matrix is needed to run the GREML models below.

```{r}
run_command(glue("{gcta} --bfile dgrp2_QC_all_lines",
                 " --make-grm",
                 " --out gcta_GRM"), path = "/gwas_data/derived/")
```

### Run univariate GREML models to find heritabilites

The below uses the `gcta64` command --reml to estimate the genetic variance (V(G)), i.e. the variance explained by the GRM), environmental variance (V(e)), their sum (Vp), and the so-called “SNP heritability” (V(G)/Vp).

```{r}

do_univariate_greml <- function(trait1){

  # First make 'focal_data', a 2-column data frame with the line and the focal phenotype value
  focal_data <- line_mean_phenotypes %>%
    select(line, !! trait1)
  names(focal_data)[2] <- c("focal_phenotype")
  
  # Write a file with the line and phenotype data called phenotype.txt, for gcta64
  pheno_data <- focal_data %>%
    # filter(!(line %in% lines_to_prune)) %>% 
    mutate(line_copy = line) %>%
    select(line, line_copy, focal_phenotype) %>%
    as.matrix()
  
  pheno_data %>%
    write.table(row.names = FALSE, col.names = FALSE,
                file = "gwas_data/derived/phenotype.txt", quote = FALSE)
  
  console_output <- suppressWarnings(capture.output(
    run_command(glue("{gcta}  --reml --reml-maxit 10000 --grm gcta_GRM  --pheno phenotype.txt  --out delete_me"), 
                path = "/gwas_data/derived/")))
  
  errors <- console_output[str_detect(console_output, "[Ee]rror")]
  if(length(errors) == 0) errors <- NA
  if(length(errors) > 1) errors <- paste(errors, sep = "; ")

  rbind(read.delim("gwas_data/derived/delete_me.hsq", sep = "\t") , 
        data.frame(Source = "Errors", Variance = errors, SE = NA)) %>% 
    mutate(Trait = trait1) %>% select(Trait, everything())
}

# Now run the model for every trait and create a large tibble using map_dfr

if(!file.exists("gwas_data/variance_components.csv")){
  variance_components <-
    map_dfr(traits_for_gwas, do_univariate_greml) %>% 
    filter(Source %in% c("V(G)", "V(e)", "Vp", "V(G)/Vp")) %>% 
    mutate(Variance = as.numeric(Variance)) %>% 
    rename(Parameter = Source) %>% 
    as_tibble()
  write_csv(variance_components, file = "gwas_data/variance_components.csv")
} else variance_components <- read_csv("gwas_data/variance_components.csv")

```

Currently two phenotypes call errors when we run the univariate GREML model: `paraquat.resistance.2021.f` and `starvation.resistance.high.carb.diet.m`. The error is _Error: the information matrix is not invertible._

For now we simply don't run models for these traits, so there are no variance estimates for these traits.

### Visualise heritability

```{r}

trait_info <- read_csv("data/all.dgrp.phenos_unscaled.csv") %>%
  mutate(Trait = str_replace_all(Trait, "[/]", "_")) %>% 
  select(-trait_value) %>% 
  distinct(Trait, Sex, `Trait guild`, `Trait description`, Reference)

variance_components <- left_join(variance_components, trait_info)

# wrangle so that each variance component is its own column

variance_components_wide <- 
  variance_components %>% 
  pivot_wider(names_from = Parameter,
              values_from = c(Variance, SE)) %>% 
  mutate(across(where(is.numeric), round, 2)) %>% 
  rename(`V(G)` = `Variance_V(G)`,
         `V(e)` = `Variance_V(e)`,
         `Vp` = `Variance_Vp`,
         `V(G)/Vp` = `Variance_V(G)/Vp`)


write_csv(variance_components_wide, file = "data/SNP_heritability.csv")
  
  
# Below we remove the suffix that denotes the sex the trait was measured in and replace the level 'both' with 'pooled'

SNP_heritability <-
  variance_components_wide %>% 
  mutate(Sex = recode(Sex, Both = "Pooled"),
         Trait = case_when(
           str_ends(Trait, ".f") | str_ends(Trait, ".m")  ~ gsub('.{2}$', '', Trait))) %>% 
  select(Trait, Sex, everything()) %>%  # reorder the columns for the table %>% 
  arrange(Trait)


write_csv(SNP_heritability, file = "data/SNP_heritability_cleaned.csv")



# create an interactive table

my_data_table <- function(df){
  datatable(
    df, rownames=FALSE,
    autoHideNavigation = TRUE,
    extensions = c("Scroller",  "Buttons"),
    options = list(
      dom = 'Bfrtip',
      deferRender=TRUE,
      scrollX=TRUE, scrollY=800,
      scrollCollapse=TRUE,
      buttons =
        list('pageLength', 'colvis', 'csv', list(
          extend = 'pdf',
          pageSize = 'A4',
          orientation = 'landscape',
          filename = 'SNP_heritability')),
      pageLength = 1000
    )
  )
}

my_data_table(SNP_heritability)

```

```{r}
p1 <-
  SNP_heritability %>% 
  ggplot(aes(x = `V(G)/Vp`)) + 
  geom_histogram(fill = met.brewer("Hokusai2")[3]) +
  labs(y = "Number of traits") +
  theme_minimal()

p1


```


## Use `GCTA` to estimate genetic covariance between traits and fitness

Here we use the `gcta64` command `--reml-bivar` to estimate the genetic covariance `(C(G)_tr12)`, i.e. the co-variance explained by the known SNPs in the DGRP), and the genetic correlation (`rG`), in addition to the univariate parameters for trait 1 and trait 2 mentioned above. Note that for traits where the genetic variance is estimated to be low, the model cannot reliably estimate genetic covariance or correlations.

We use `rG` to calculate the Robertson-Price identity, which estimates the response to selection for a trait (or the selection differential... check).

### Female fitness 
 
```{r} 
# wrangle data so that it is specific to the sex the trait was measured for i.e. female, male or pooled sexes

female_line_means <- line_mean_phenotypes %>%
  select(line, ends_with(".f"))

female_trait_list <- line_mean_phenotypes %>% 
  select(ends_with(".f")) %>% 
  select(-fitness.early.f) %>% 
  colnames()

# build the function to run the bivariate GREML  

do_bivar_greml_f <- function(trait1, trait2){

  # First make 'focal_data', a 2-column data frame with the line and the focal phenotype value
  focal_data <- female_line_means %>%
    select(line, !! trait1, !! trait2)
  names(focal_data)[2:3] <- c("focal_pheno1", "focal_pheno2")
  
  # Write a file with the line and phenotype data called phenotype.txt, for gcta64
  pheno_data <- focal_data %>%
    # filter(!(line %in% lines_to_prune)) %>% 
    mutate(line_copy = line) %>%
    select(line, line_copy, focal_pheno1, focal_pheno2) %>%
    as.matrix()
  
  pheno_data %>%
    write.table(row.names = FALSE, col.names = FALSE,
                file = "gwas_data/derived/phenotype_pair.txt", quote = FALSE)
  
  # Note that the warnings come when gcta could not fit the model properly. I save these errors especially, so no need to print them.
  console_output <- suppressWarnings(capture.output(
    run_command(glue("{gcta}  --reml-bivar --reml-bivar-lrt-rg 0 --reml-maxit 10000 --grm gcta_GRM  --pheno phenotype_pair.txt  --out delete_me"), 
                path = "/gwas_data/derived/")))
  
  errors <- console_output[str_detect(console_output, "[Ee]rror")]
  if(length(errors) == 0) errors <- NA
  if(length(errors) > 1) errors <- paste(errors, sep = "; ")
  traits <- sort(c(trait1, trait2))
  
  rbind(read.delim("gwas_data/derived/delete_me.hsq", sep = "\t") , 
        data.frame(Source = "Errors", Variance = errors, SE = NA)) %>% 
    mutate(`Trait 1` = traits[1], `Trait 2` = traits[2]) %>% 
    select(`Trait 1`, `Trait 2`, everything())
}

# run for all traits - this works, except I need to rewrite the function for each level of sex (done below)

if(!file.exists("data/SNP_selection.csv")){
  female_results <- map2_dfr("fitness.early.f", female_trait_list, do_bivar_greml_f)
  # unlink deletes files
  unlink(c("data/derived/delete_me.log", "data/derived/phenotype_pair.txt", "data/derived/delete_me.hsq"))
  unlink(list.files("data/derived/", pattern = "gcta_GRM", full.names = TRUE))
  # clean the data up
  female_results_concise <-
    female_results %>% 
    as_tibble() %>% 
    filter(Source %in% c("V(G)_tr1", "V(e)_tr1", "Vp_tr1", "V(G)/Vp_tr1",
                         "V(G)_tr2", "V(e)_tr2", "Vp_tr2", "V(G)/Vp_tr2",
                         "C(G)_tr12", "rG")) %>% 
    mutate(Variance = as.numeric(Variance),
           Fitness_sex = "Female") %>% 
    rename(Parameter = Source)
} else "Already calculated"

```

### Male fitness

```{r}
# wrangle data so that it is specific to the sex the trait was measured for i.e. female, male or pooled sexes

male_line_means <- line_mean_phenotypes %>%
  select(line, ends_with(".m"))

male_trait_list <- line_mean_phenotypes %>% 
  select(ends_with(".m")) %>% 
  select(-fitness.early.m) %>% 
  colnames()

# build the function to run the bivariate GREML  

do_bivar_greml_m <- function(trait1, trait2){

  # First make 'focal_data', a 2-column data frame with the line and the focal phenotype value
  focal_data <- male_line_means %>%
    select(line, !! trait1, !! trait2)
  names(focal_data)[2:3] <- c("focal_pheno1", "focal_pheno2")
  
  # Write a file with the line and phenotype data called phenotype.txt, for gcta64
  pheno_data <- focal_data %>%
    # filter(!(line %in% lines_to_prune)) %>% 
    mutate(line_copy = line) %>%
    select(line, line_copy, focal_pheno1, focal_pheno2) %>%
    as.matrix()
  
  pheno_data %>%
    write.table(row.names = FALSE, col.names = FALSE,
                file = "gwas_data/derived/phenotype_pair.txt", quote = FALSE)
  
  # Note that the warnings come when gcta could not fit the model properly. I save these errors especially, so no need to print them.
  console_output <- suppressWarnings(capture.output(
    run_command(glue("{gcta}  --reml-bivar --reml-bivar-lrt-rg 0 --reml-maxit 10000 --grm gcta_GRM  --pheno phenotype_pair.txt  --out delete_me"), 
                path = "/gwas_data/derived/")))
  
  errors <- console_output[str_detect(console_output, "[Ee]rror")]
  if(length(errors) == 0) errors <- NA
  if(length(errors) > 1) errors <- paste(errors, sep = "; ")
  traits <- sort(c(trait1, trait2))
  
  rbind(read.delim("gwas_data/derived/delete_me.hsq", sep = "\t") , 
        data.frame(Source = "Errors", Variance = errors, SE = NA)) %>% 
    mutate(`Trait 1` = traits[1], `Trait 2` = traits[2]) %>% 
    select(`Trait 1`, `Trait 2`, everything())
}

# run for all traits - this works, except I need to rewrite the function for each level of sex (done below)

if(!file.exists("data/SNP_selection.csv")){
  male_results <- map2_dfr("fitness.early.m", male_trait_list, do_bivar_greml_m)
  # unlink deletes files
  unlink(c("data/derived/delete_me.log", "data/derived/phenotype_pair.txt", "data/derived/delete_me.hsq"))
  unlink(list.files("data/derived/", pattern = "gcta_GRM", full.names = TRUE))
  # clean the data up
  male_results_concise <-
    male_results %>% 
    as_tibble() %>% 
    filter(Source %in% c("V(G)_tr1", "V(e)_tr1", "Vp_tr1", "V(G)/Vp_tr1",
                         "V(G)_tr2", "V(e)_tr2", "Vp_tr2", "V(G)/Vp_tr2",
                         "C(G)_tr12", "rG")) %>% 
    mutate(Variance = as.numeric(Variance),
           Fitness_sex = "Male") %>% 
    rename(Parameter = Source)
} else "Already calculated"

```

### Data where the sexes were pooled

```{r}

# wrangle data so that it is specific to the sex the trait was measured for i.e. female, male or pooled sexes

pooled_line_means <- line_mean_phenotypes %>%
  select(line, fitness.early.f, fitness.early.m, !ends_with(c(".f", ".m")))

pooled_trait_list <- line_mean_phenotypes %>% 
  select(!ends_with(c(".f", ".m"))) %>%  
  select(-line) %>% 
  colnames()

# build the function to run the bivariate GREML  

do_bivar_greml_p <- function(trait1, trait2){

  # First make 'focal_data', a 2-column data frame with the line and the focal phenotype value
  focal_data <- pooled_line_means %>%
    select(line, !! trait1, !! trait2)
  names(focal_data)[2:3] <- c("focal_pheno1", "focal_pheno2")
  
  # Write a file with the line and phenotype data called phenotype.txt, for gcta64
  pheno_data <- focal_data %>%
    # filter(!(line %in% lines_to_prune)) %>% 
    mutate(line_copy = line) %>%
    select(line, line_copy, focal_pheno1, focal_pheno2) %>%
    as.matrix()
  
  pheno_data %>%
    write.table(row.names = FALSE, col.names = FALSE,
                file = "gwas_data/derived/phenotype_pair.txt", quote = FALSE)
  
  # Note that the warnings come when gcta could not fit the model properly. I save these errors especially, so no need to print them.
  console_output <- suppressWarnings(capture.output(
    run_command(glue("{gcta}  --reml-bivar --reml-bivar-lrt-rg 0 --reml-maxit 10000 --grm gcta_GRM  --pheno phenotype_pair.txt  --out delete_me"), 
                path = "/gwas_data/derived/")))
  
  errors <- console_output[str_detect(console_output, "[Ee]rror")]
  if(length(errors) == 0) errors <- NA
  if(length(errors) > 1) errors <- paste(errors, sep = "; ")
  traits <- sort(c(trait1, trait2))
  
  rbind(read.delim("gwas_data/derived/delete_me.hsq", sep = "\t") , 
        data.frame(Source = "Errors", Variance = errors, SE = NA)) %>% 
    mutate(`Trait 1` = traits[1], `Trait 2` = traits[2]) %>% 
    select(`Trait 1`, `Trait 2`, everything())
}

# run for all traits - this works, except I need to rewrite the function for each level of sex (done below)

if(!file.exists("data/SNP_selection.csv")){
  pooled_female <- map2_dfr("fitness.early.f", pooled_trait_list, do_bivar_greml_p)
  # unlink deletes files
  unlink(c("data/derived/delete_me.log", "data/derived/phenotype_pair.txt", "data/derived/delete_me.hsq"))
  unlink(list.files("data/derived/", pattern = "gcta_GRM", full.names = TRUE))
  pooled_male <- map2_dfr("fitness.early.m", pooled_trait_list, do_bivar_greml_p)
  # unlink deletes files
  unlink(c("data/derived/delete_me.log", "data/derived/phenotype_pair.txt", "data/derived/delete_me.hsq"))
  unlink(list.files("data/derived/", pattern = "gcta_GRM", full.names = TRUE))
  # clean the data up
  pooled_results <- rbind(pooled_female, pooled_male)
  
  pooled_results_concise <-
    pooled_results %>% 
    as_tibble() %>% 
    filter(Source %in% c("V(G)_tr1", "V(e)_tr1", "Vp_tr1", "V(G)/Vp_tr1",
                         "V(G)_tr2", "V(e)_tr2", "Vp_tr2", "V(G)/Vp_tr2",
                         "C(G)_tr12", "rG")) %>% 
    mutate(Variance = as.numeric(Variance),
           Fitness_sex = if_else(`Trait 2` == "fitness.early.f", "Female", "Male")) %>% 
    rename(Parameter = Source)
} else "Already calculated"

```

### Combine results and save

Note that when we combine the `SNP_heritability` tibble with the `SNP_selection` tibble we will get some duplicate columns. This is because the bivariate GREML calculates everything that the univariate GREML does. I'll check if they are identical and delete redundant columns if so. I get the feeling that they'll be slightly different because the bivariate GREML only uses the lines that have been measured for both traits; generally a smaller subset of those a trait was measured across.

```{r}

if(!file.exists("data/SNP_selection.csv")){
  SNP_selection_response_estimates <-
    rbind(female_results_concise, male_results_concise, pooled_results_concise) %>% 
    # create a common name so that we can join the meta data
    rename(Trait = `Trait 1`)
  write_csv(SNP_selection_response_estimates, file = "data/SNP_selection.csv")
} else SNP_selection_response_estimates <- read_csv("data/SNP_selection.csv")

# join the meta data

SNP_selection_response_estimates <- left_join(SNP_selection_response_estimates, trait_info)

# wrangle so that each variance component is its own column

SNP_selection_response_estimates_wide <-
  SNP_selection_response_estimates %>% 
  pivot_wider(names_from = Parameter, values_from = c(Variance, SE)) %>% 
  unnest(cols = c(`Variance_V(G)_tr1`, `Variance_V(G)_tr2`, `Variance_C(G)_tr12`, 
                  `Variance_V(e)_tr1`, `Variance_V(e)_tr2`, Variance_Vp_tr1, 
                  Variance_Vp_tr2, `Variance_V(G)/Vp_tr1`, `Variance_V(G)/Vp_tr2`, 
                  Variance_rG, `SE_V(G)_tr1`, `SE_V(G)_tr2`, `SE_C(G)_tr12`, 
                  `SE_V(e)_tr1`, `SE_V(e)_tr2`, SE_Vp_tr1, SE_Vp_tr2, `SE_V(G)/Vp_tr1`, 
                  `SE_V(G)/Vp_tr2`, SE_rG))

# this part here is to make sure fitness is in the same trait column throughout the tibble. Trait 1 is determined alphabetically, so it's sometimes fitness and sometimes is not. We fix that below 

a <- 
  SNP_selection_response_estimates_wide %>% 
  filter(Trait == "fitness.early.f" | Trait == "fitness.early.m") %>% 
  rename(`Trait 1` = Trait,
         # rename to more specific names. No name switching needed for this subset
         `Variance_V(G)_trait` = `Variance_V(G)_tr2`,
         `Variance_V(G)_fitness` = `Variance_V(G)_tr1`,
         `Variance_V(e)_trait` = `Variance_V(e)_tr2`,
         `Variance_V(e)_fitness` = `Variance_V(e)_tr1`,
         Variance_Vp_trait = Variance_Vp_tr2,
         Variance_Vp_fitness = Variance_Vp_tr1,
         `Variance_V(G)/Vp_trait` = `Variance_V(G)/Vp_tr2`,
         `Variance_V(G)/Vp_fitness` = `Variance_V(G)/Vp_tr1`,
         `SE_V(G)_trait` = `SE_V(G)_tr2`,
         `SE_V(G)_fitness` = `SE_V(G)_tr1`,
         `SE_V(e)_trait` = `SE_V(e)_tr2`,
         `SE_V(e)_fitness` = `SE_V(e)_tr1`,
         SE_Vp_trait = SE_Vp_tr2,
         SE_Vp_fitness = SE_Vp_tr1,
         `SE_V(G)/Vp_trait` = `SE_V(G)/Vp_tr2`,
         `SE_V(G)/Vp_fitness` = `SE_V(G)/Vp_tr1`)

b <-
  SNP_selection_response_estimates_wide %>% 
  filter(`Trait 2` == "fitness.early.f" | `Trait 2` == "fitness.early.m") %>% 
  rename(`Trait 1` = `Trait 2`,
         `Trait 2` = Trait,
         # switch the names so that they match the order of trait1 and trait2
         `Variance_V(G)_trait` = `Variance_V(G)_tr1`,
         `Variance_V(G)_fitness` = `Variance_V(G)_tr2`,
         `Variance_V(e)_trait` = `Variance_V(e)_tr1`,
         `Variance_V(e)_fitness` = `Variance_V(e)_tr2`,
         Variance_Vp_trait = Variance_Vp_tr1,
         Variance_Vp_fitness = Variance_Vp_tr2,
         `Variance_V(G)/Vp_trait` = `Variance_V(G)/Vp_tr1`,
         `Variance_V(G)/Vp_fitness` = `Variance_V(G)/Vp_tr2`,
         `SE_V(G)_trait` = `SE_V(G)_tr1`,
         `SE_V(G)_fitness` = `SE_V(G)_tr2`,
         `SE_V(e)_trait` = `SE_V(e)_tr1`,
         `SE_V(e)_fitness` = `SE_V(e)_tr2`,
         SE_Vp_trait = SE_Vp_tr1,
         SE_Vp_fitness = SE_Vp_tr2,
         `SE_V(G)/Vp_trait` = `SE_V(G)/Vp_tr1`,
         `SE_V(G)/Vp_fitness` = `SE_V(G)/Vp_tr2`)

SNP_selection_response_estimates_wide <-
  rbind(a, b) %>% 
  rename(Fitness_component = `Trait 1`,
         Trait = `Trait 2`)


SNP_selection_response_estimates_wide <-
  left_join(SNP_selection_response_estimates_wide, trait_info) %>% 
  select(Trait, Fitness_component, Fitness_sex, Sex, `Trait guild`, `Trait description`, Reference, everything()) %>% 
  # Trait_sex shows which sex the trait was measured in, while Fitness_sex shows the sex that `(C(G)_tr12)` and `rG` were calculated for
  # These are generally the same, with the exception of traits that were measured with the sexes pooled (commonly larval traits)
  # For these pooled traits we ran the bivariate GREML twice - once for female fitness and again for male fitness
  rename(Trait_sex = Sex)


```

```{r}
all_SNP_data <-
  left_join(SNP_selection_response_estimates_wide, SNP_heritability)
```

$~$

## Is heritability sex-specific?


```{r}
full_SNP_heritability_plot <- 
  SNP_selection_response_estimates_wide %>% 
  filter(`Variance_C(G)_tr12` > -1 & `Variance_V(G)/Vp_trait` > -1 & `Variance_C(G)_tr12` < 1 & `Variance_V(G)/Vp_trait` < 1) %>%
  
  ggplot(aes(x = Fitness_sex, y = `Variance_V(G)/Vp_trait`, fill = Fitness_sex)) +
  geom_jitter(shape = 21, size = 3, alpha = 0.75) +
  scale_fill_manual(values = met.brewer("VanGogh2", 2)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.8, fill = "white") +
  labs(title = "All traits measured",
       x = "Sex",
       y = "SNP heritability") +
  theme_bw() +
  theme(legend.position = "none")

  
```


```{r}
data <- SNP_selection_response_estimates_wide %>% 
  mutate(Trait = case_when(
    str_ends(Trait, ".f") | str_ends(Trait, ".m")  ~ gsub('.{2}$', '', Trait)))

a <- data %>% 
  filter(Fitness_sex == "Male")
 

b <- data %>% 
  filter(Fitness_sex == "Female")

# there are 264 traits for which we have measures in both sexes

SNP_both_sexes <- inner_join(a, b, by = "Trait") %>% 
  filter(Trait != "NA")

subset_SNP_heritability_plot <- 
  SNP_both_sexes %>% 
  pivot_longer(cols = c(`Variance_V(G)/Vp_trait.x`, `Variance_V(G)/Vp_trait.y`), names_to = "Sex", values_to = "SNP_heritability") %>% 
  mutate(Sex = if_else(str_ends(Sex, ".x"), "Male", "Female")) %>% 
  select(-c(Fitness_sex.x, Fitness_sex.y)) %>% 
  filter(SNP_heritability > -1 & SNP_heritability < 1) %>%
  
  ggplot(aes(x = Sex, y = SNP_heritability, fill = Sex)) +
  geom_jitter(shape = 21, size = 3, alpha = 0.75) +
  scale_fill_manual(values = met.brewer("VanGogh2", 2)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.8, fill = "white") +
  labs(title = "Traits measured in both sexes",
       x = "Sex",
       y = NULL) +
  theme_bw() +
  theme(legend.position = "none")

full_SNP_heritability_plot + subset_SNP_heritability_plot
```



## Testing general predictions of Fisher's theorem 

$~$

**1. Do traits correlated with fitness have lower heritabilities?**

Following Fisher's theorem and assuming a constant environment, in the absence of new variation, selection is expected to erode all additive genetic variation for fitness. One implication of this expectation is that traits that are correlated with fitness should show reduced additive genetic variance relative to those that are not correlated with fitness _(Robertson 1955b)_. 

There are previous studies that have compared fitness related traits i.e. those implicated with reproduction, with morphological traits, which are assumed to have lesser relationships with fitness.

- The problem is associating these traits with _lifetime) fitness. This shouldn't be such a problem for us...

- Multiple studies looking at this corollary use phenotypic correlations with fitness rather than genetic. Once again, this is no problem for us...

- Even if there is lots of additive genetic variance for a trait and it is correlated with fitness, if fitness has a high environmental variance component the strength of selection on the trait will be weak.

_Houle (1992)_ looks like an important meta-analysis type work


```{r}
# some plots

a_plot <-
  SNP_selection_response_estimates_wide %>% 
  filter(`Variance_C(G)_tr12` > -1 & Variance_rG > -1 & `Variance_C(G)_tr12` < 1 & Variance_rG < 1) %>% 
         #`Trait guild` != "Metabolome" & `Trait guild` != "CHC", `Trait guild` != "Microbiome") %>% 
  ggplot(aes(x = `Variance_C(G)_tr12`, y = Variance_rG, fill = Fitness_sex)) +
  geom_point(shape = 21, size = 3, alpha = 0.5) +
  scale_fill_manual(values = met.brewer("VanGogh2", 2)) +
  geom_abline(intercept = 0, slope = 1, linetype = 2) +
  labs(x = "SNP covariance between fitness and trait",
       y = "Genetic correlation between fitness\nand trait (Rg)",
       fill = "Sex") +
  theme_bw() 


b_plot <-
  SNP_selection_response_estimates_wide %>% 
  filter(`Variance_C(G)_tr12` > -1 & `Variance_V(G)/Vp_trait` > -1 & `Variance_C(G)_tr12` < 1 & `Variance_V(G)/Vp_trait` < 1) %>% 
  #`Trait guild` != "Metabolome" & `Trait guild` != "CHC", `Trait guild` != "Microbiome") %>% 
  mutate(`Variance_C(G)_tr12` = abs(`Variance_C(G)_tr12`)) %>% 
  ggplot(aes(x = `Variance_C(G)_tr12`, y = `Variance_V(G)/Vp_trait`, fill = Fitness_sex)) +
  #geom_hex()
  geom_point(shape = 21, size = 3, alpha = 0.5) +
  geom_smooth(method = "lm", colour = "black") +
  scale_fill_manual(values = met.brewer("VanGogh2", 2)) +
  labs(x = "SNP covariance between fitness\n and trait",
       y = "SNP heritability") +
  theme_bw() +
  theme(legend.position = "none")

c_plot <-
  SNP_selection_response_estimates_wide %>% 
  filter(Variance_rG > -1 & `Variance_V(G)/Vp_trait` > -1 & Variance_rG < 1 & `Variance_V(G)/Vp_trait` < 1) %>% 
  #`Trait guild` != "Metabolome" & `Trait guild` != "CHC", `Trait guild` != "Microbiome") %>% 
  mutate(Variance_rG = abs(Variance_rG)) %>% 
  ggplot(aes(x = Variance_rG, y = `Variance_V(G)/Vp_trait`, fill = Fitness_sex)) +
  #geom_hex() +
  geom_point(shape = 21, size = 3, alpha = 0.5) +
  geom_smooth(method = "lm", colour = "black") +
  scale_fill_manual(values = met.brewer("VanGogh2", 2)) +
  labs(x = "Genetic correlation (Rg)\n between fitness and trait",
       y = "SNP heritability") +
  theme_bw() +
  theme(legend.position = "none")

a_plot / (b_plot + c_plot)
```

Figure X: **a** the relationship between the genetic correlation between fitness and trait (Rg) and the SNP covariance between fitness and trait. Both of these metrics could **potentially** be used to estimate the response to selection experienced by a trait (or the selection differential, not sure yet), following the Robertson-Price identity. **b** and **c** show the relationship between these metrics and heritability.

$~$

**2. Do traits correlated with fitness have higher levels of both additive and residual variance**

Note that heritability can also decline with an increase in environmental variance and no decrease (or an increase) in genetic variance. If true, this makes a simple regression with heritability and fitness misleading.

Using evolvability as his metric, _Houle (1992)_ found that characters assumed to be closely related to fitness have higher evolvabilities than do trait with looser relationships to fitness. This suggests that the negative relationship between fitness and heritability is due to increased environmental variance in fitness related traits, not a decrease in additive genetic variance. 

$~$

## Sexually antagonstic selection

$~$

We have found selection differentials **(or responses to selection)** acting on traits in females and males. For those traits where we have a measure of both, we can quantify sexually antagonistic selection by finding traits that have selection operating in opposite directions, depending on which sex the trait is expressed in. 

[Innocenti and Morrow (2011)]( https://doi.org/10.1111/j.1558-5646.2010.01021.x) present an index for measuring the intensity for sexually antagonistic selection:

$I = \frac{\beta_M \beta_F}{\sqrt{(\beta_M^{2} + \beta_F^{2})/2}}$


Where $B_F$ and $B_M$ are the standardised selection gradients for females and males respectively.

From **Innocenti and Morrow**:

This index is positive when selection is concordant in the two sexes, negative when antagonistic in the two sexes, and is zero when selection is absent in one sex (note that it will miss conflict that occurs when strong stabilizing selection is present in that sex). Finally it is proportional to the absolute intensity of selection.

Additionally, it has the desirable properties of being symmetrical and normally distributed for a random set of normally distributed $B_F$, $B_M$. |I| is also always included in the interval between the absolute values of the selection gradient in the two sexes, and it coincides with them when $B_F$ = $B_M$. As a potential drawback, it should be noted that such quantity is not defined when $B_F$ = $B_M$ = 0, even though it makes little sense to estimate how concordant or antagonistic directional selection is when it is absent in both sexes.


$~$

### Covariance

```{r}


# create our sexual antagonism tibble by joining the female and male tibbles by traits they both share

SNP_antagonism <- SNP_both_sexes %>% 
  rename(Trait_covariance_male_fitness = `Variance_C(G)_tr12.x`,
         Trait_covariance_female_fitness = `Variance_C(G)_tr12.y`) %>% 
  select(-c(Fitness_sex.x, Fitness_sex.y)) %>% 
  # calculate the innocenti and morrow index
  mutate(Selection_index = Trait_covariance_female_fitness * Trait_covariance_male_fitness / sqrt(((Trait_covariance_female_fitness)^2 + (Trait_covariance_male_fitness)^2)/2)) %>% 
  # get rid of traits with unrealistic covariances
  filter(Trait_covariance_male_fitness > -5,
         Trait_covariance_female_fitness < 2 & Trait_covariance_female_fitness > -5)

# plot the data

sexual_concordance_plot <-
  ggplot(data = SNP_antagonism, aes(x = Trait_covariance_female_fitness, y = Trait_covariance_male_fitness, fill = Selection_index)) +
  #geom_hex(bins = 20) +
  geom_point(shape = 21, alpha = 0.85, size = 4, show.legend = TRUE) + #width = 0.003, height = 0.003) +
  #geom_point(shape = 21, alpha = 0.9, size = 5, show.legend = FALSE) +
  geom_smooth(method = 'lm', color='black') +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  #coord_cartesian(xlim = c(-0.35, 0.35), ylim = c(-0.4, 0.4)) +
  scale_fill_carto_c(palette = "Geyser",
                     breaks=c(-0.25, -0.125, 0, 0.125, 0.25),
                     limits=c(-0.25,0.25)) +
  labs(x = "Female selection differential",
       y = "Male selection differential",
       fill = "Concordance\n of selection\n") +
  theme_bw() +
  theme(panel.border= element_blank(),
        axis.line=element_line(),
        text = element_text(size=14),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

sexual_concordance_plot


```

### Genetic correlation

```{r}
data <- SNP_selection_response_estimates_wide %>% 
  select(Trait, Fitness_sex, Trait_sex, Variance_rG) %>% 
  mutate(Trait = case_when(
    str_ends(Trait, ".f") | str_ends(Trait, ".m")  ~ gsub('.{2}$', '', Trait)))

a <- data %>% 
  filter(Fitness_sex == "Male") %>% 
  select(-Trait_sex) 
 

b <- data %>% 
  filter(Fitness_sex == "Female") %>% 
  select(-Trait_sex)


# create our sexual antagonism tibble by joining the female and male tibbles by traits they both share

SNP_antagonism_rg <- inner_join(a, b, by = "Trait") %>% 
  rename(Trait_rg_male_fitness = Variance_rG.x,
         Trait_rg_female_fitness = Variance_rG.y) %>% 
  select(-c(Fitness_sex.x, Fitness_sex.y)) %>% 
  # calculate the innocenti and morrow index
  mutate(Selection_index = Trait_rg_female_fitness * Trait_rg_male_fitness / sqrt(((Trait_rg_female_fitness)^2 + (Trait_rg_male_fitness)^2)/2)) %>% 
  # get rid of traits with unrealistic covariances
  filter(Trait != "NA",
         Trait_rg_male_fitness > -1 & Trait_rg_male_fitness < 1,
         Trait_rg_female_fitness < 1 & Trait_rg_female_fitness > -1)

sexual_concordance_plot_rg <-
  ggplot(data = SNP_antagonism_rg, aes(x = Trait_rg_female_fitness, y = Trait_rg_male_fitness, fill = Selection_index)) +
  #geom_hex(bins = 20) +
  geom_point(shape = 21, alpha = 0.75, size = 4, show.legend = TRUE) + #width = 0.003, height = 0.003) +
  #geom_point(shape = 21, alpha = 0.9, size = 5, show.legend = FALSE) +
  geom_smooth(method = 'lm', color='black') +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_vline(xintercept = 0, linetype = 2) +
  #coord_cartesian(xlim = c(-0.4, 0.4), ylim = c(-0.4, 0.4)) +
  scale_fill_carto_c(palette = "Geyser",
                     breaks=c(-1, -0.5, 0, 0.5, 1),
                     limits=c(-1, 1)) +
  labs(x = "Female selection differential (Rg)",
       y = "Male selection differential (Rg)",
       fill = "Concordance\n of selection\n") +
  theme_bw() +
  theme(panel.border= element_blank(),
        axis.line=element_line(),
        text = element_text(size=14),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14))

sexual_concordance_plot_rg
```

