---
title: "Correlations in SNP effect sizes from GWAS"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Load packages and custom functions

```{r}
library(tidyverse)
library(DT)
library(glue)
library(pheatmap)
library(RColorBrewer)
library(coop) # for covar, a faster version of R's cov function

files <- list.files("gwas_data/derived/gwas_results", pattern = "tsv.gz", full.names = TRUE)
traits <- str_split(files, "/") %>% map_chr(~ .x[4]) %>% str_remove_all(".tsv.gz")


trait_guilds <- read_csv("data/all.dgrp.phenos_unscaled.csv") %>% 
  select(Trait, `Trait guild`) %>% distinct() %>% 
  mutate(`Trait guild` = replace(`Trait guild`, is.na(`Trait guild`), "Other"),
         `Trait guild` = replace(`Trait guild`, `Trait guild` == "?", "Other"))

guilds <- unique(trait_guilds$`Trait guild`)
```

```{r}
my_data_table <- function(df){
  datatable(
    df, rownames=FALSE,
    autoHideNavigation = TRUE,
    extensions = c("Scroller",  "Buttons"),
    options = list(
      dom = 'Bfrtip',
      deferRender=TRUE,
      scrollX=TRUE, scrollY=800,
      scrollCollapse=TRUE,
      buttons =
        list('pageLength', 'colvis', 'csv', list(
          extend = 'pdf',
          pageSize = 'A4',
          orientation = 'landscape',
          filename = 'selection_data')),
      pageLength = 905
    )
  )
}

# Loads all the SNP effect size data for all the traits in the focal guild (data is too big to load all traits at once)
# It also additionally loads the SNP effects for male and female early-life fitness 
load_guild <- function(focal_guild){
  traits_in_guild <- trait_guilds %>% filter(`Trait guild` == focal_guild) %>% pull(Trait)
  focal_files <- unique(c(files[traits %in% traits_in_guild], files[str_detect(files, "fitness")]))
  focal_traits <- unique(c(traits[traits %in% traits_in_guild], traits[str_detect(traits, "fitness")]))
  
  lapply(1:length(focal_files), function(i){
    trait <- focal_traits[i]
    dat <- read_tsv(files[i]) %>% select(SNP, BETA)
    names(dat)[2] <- trait
    dat
  }) %>% reduce(full_join, by = "SNP")
}

# Make a correlation matrix of SNP effects, using coop::covar and then cov2cor
make_corr_matrix <- function(snp_effect_data){
  snp_effect_data <- snp_effect_data[!(names(snp_effect_data) == "SNP")]
  corr_matrix <- cov2cor(covar(snp_effect_data, use = "complete.obs"))
  corr_matrix[corr_matrix > 0.999] <- NA # some self-correlations were not exactly 1, due to floating point numbers
  colnames(corr_matrix) <- names(snp_effect_data)
  rownames(corr_matrix) <- names(snp_effect_data)
  corr_matrix
}

# Make nice pheatmaps
make_heatmap <- function(corr_matrix, cuts = 6){
  paletteLength <- 50
  myColor <- colorRampPalette(c(brewer.pal(9, "Purples")[7], "white", brewer.pal(9, "Oranges")[7]))(paletteLength)
  myBreaks <- c(seq(min(corr_matrix, na.rm=T), 0, length.out=ceiling(paletteLength/2) + 1), 
                seq(max(corr_matrix, na.rm=T)/paletteLength, max(corr_matrix, na.rm=T), length.out=floor(paletteLength/2)))
  
  if(is.na(cuts)){
    pheatmap(corr_matrix, show_rownames = TRUE, col = myColor,
             cutree_rows = cuts, cutree_cols = cuts)
  }
  else {
    pheatmap(corr_matrix, show_rownames = TRUE, col = myColor)
  }
}

# Make a neat table showing all the correlations for a given set of data
make_correlation_table <- function(snp_effect_data){
  all_corrs <- snp_effect_data %>% 
    make_corr_matrix() %>% as.data.frame() %>% 
    rownames_to_column("trait1") %>% 
    gather(trait2, corr, -trait1) %>% 
    filter(!is.na(corr))
  
  for(i in 1:nrow(all_corrs)){
    all_corrs[i, 1:2] <- sort(all_corrs[i, 1:2]) 
  }
  
  all_corrs %>% 
    mutate(corr = round(corr, 4)) %>% 
    distinct() %>% 
    as_tibble() %>% 
    arrange(-(corr)) %>% 
    mutate(trait1_sex = substr(trait1, nchar(trait1)-1, nchar(trait1)),
           trait2_sex = substr(trait2, nchar(trait2)-1, nchar(trait2)),
           trait1 = replace(trait1, trait1_sex == ".m", substr(trait1, 1, nchar(trait1)-2)[trait1_sex == ".m"]),
           trait1 = replace(trait1, trait1_sex == ".f", substr(trait1, 1, nchar(trait1)-2)[trait1_sex == ".f"]),
           trait2 = replace(trait2, trait2_sex == ".m", substr(trait2, 1, nchar(trait2)-2)[trait2_sex == ".m"]),
           trait2 = replace(trait2, trait2_sex == ".f", substr(trait2, 1, nchar(trait2)-2)[trait2_sex == ".f"]),
           trait1_sex = case_when(trait1_sex == ".f" ~ "female", 
                                  trait1_sex == ".m" ~ "male", TRUE ~ "both" ),
           trait2_sex = case_when(trait2_sex == ".f" ~ "female", 
                                  trait2_sex == ".m" ~ "male", TRUE ~ "both" ),
           corr_sex_type = ifelse(trait1_sex == "female" & trait2_sex == "male", "between_sex", NA),
           corr_sex_type = replace(corr_sex_type, trait2_sex == "female" & trait1_sex == "male", "between_sex"),
           corr_sex_type = replace(corr_sex_type, trait1_sex == trait2_sex, "within_sex"))
}
```

## Unclassified traits {.tabset}

### Heatmap
```{r}
snp_effect_data <- load_guild("Other") 

snp_effect_data %>% 
  make_corr_matrix() %>% make_heatmap()
```

### Fitness-trait correlations
```{r}
all_corrs <- snp_effect_data %>% make_correlation_table()

all_corrs %>% 
  filter(str_detect(trait1, "fitness") | str_detect(trait2, "fitness")) %>% 
  my_data_table()
```

### Between-sex correlations for shared traits

```{r}
all_corrs %>% 
  filter(corr_sex_type == "between_sex" &  trait1 == trait2) %>% 
  arrange(-corr) %>% 
  my_data_table()
```


## Morphological traits {.tabset}

### Heatmap
```{r}
snp_effect_data <- load_guild("Morphological") 

snp_effect_data %>% 
  make_corr_matrix() %>% make_heatmap()
```

### Fitness-trait correlations
```{r}
all_corrs <- snp_effect_data %>% make_correlation_table()

all_corrs %>% 
  filter(str_detect(trait1, "fitness") | str_detect(trait2, "fitness")) %>% 
  my_data_table()
```

### Between-sex correlations for shared traits

```{r}
all_corrs %>% 
  filter(corr_sex_type == "between_sex" &  trait1 == trait2) %>% 
  arrange(-corr) %>% 
  my_data_table()
```


## Physiological traits {.tabset}

### Heatmap
```{r}
snp_effect_data <- load_guild("Physiological") 

snp_effect_data %>% 
  make_corr_matrix() %>% make_heatmap()
```

### Fitness-trait correlations
```{r}
all_corrs <- snp_effect_data %>% make_correlation_table()

all_corrs %>% 
  filter(str_detect(trait1, "fitness") | str_detect(trait2, "fitness")) %>% 
  my_data_table()
```

### Between-sex correlations for shared traits

```{r}
all_corrs %>% 
  filter(corr_sex_type == "between_sex" &  trait1 == trait2) %>% 
  arrange(-corr) %>% 
  my_data_table()
```


## Behavioural traits {.tabset}

### Heatmap
```{r}
snp_effect_data <- load_guild("Behavioural") 

snp_effect_data %>% 
  make_corr_matrix() %>% make_heatmap(cuts = NA)
```

### Fitness-trait correlations
```{r}
all_corrs <- snp_effect_data %>% make_correlation_table()

all_corrs %>% 
  filter(str_detect(trait1, "fitness") | str_detect(trait2, "fitness")) %>% 
  my_data_table()
```

### Between-sex correlations for shared traits

```{r}
all_corrs %>% 
  filter(corr_sex_type == "between_sex" &  trait1 == trait2) %>% 
  arrange(-corr) %>% 
  my_data_table()
```



## Life history traits {.tabset}

### Heatmap
```{r}
snp_effect_data <- load_guild("Life history") 

snp_effect_data %>% 
  make_corr_matrix() %>% make_heatmap()
```

### Fitness-trait correlations
```{r}
all_corrs <- snp_effect_data %>% make_correlation_table()

all_corrs %>% 
  filter(str_detect(trait1, "fitness") | str_detect(trait2, "fitness")) %>% 
  my_data_table()
```

### Between-sex correlations for shared traits

```{r}
all_corrs %>% 
  filter(corr_sex_type == "between_sex" &  trait1 == trait2) %>% 
  arrange(-corr) %>% 
  my_data_table()
```


## Pathogen response traits {.tabset}

### Heatmap
```{r}
snp_effect_data <- load_guild("Pathogen response") 

snp_effect_data %>% 
  make_corr_matrix() %>% make_heatmap()
```

### Fitness-trait correlations
```{r}
all_corrs <- snp_effect_data %>% make_correlation_table()

all_corrs %>% 
  filter(str_detect(trait1, "fitness") | str_detect(trait2, "fitness")) %>% 
  my_data_table()
```

### Between-sex correlations for shared traits

```{r}
all_corrs %>% 
  filter(corr_sex_type == "between_sex" &  trait1 == trait2) %>% 
  arrange(-corr) %>% 
  my_data_table()
```



## Insecticide response traits {.tabset}

### Heatmap
```{r}
snp_effect_data <- load_guild("Insecticide response") 

snp_effect_data %>% 
  make_corr_matrix() %>% make_heatmap()
```

### Fitness-trait correlations
```{r}
all_corrs <- snp_effect_data %>% make_correlation_table()

all_corrs %>% 
  filter(str_detect(trait1, "fitness") | str_detect(trait2, "fitness")) %>% 
  my_data_table()
```

### Between-sex correlations for shared traits

```{r}
all_corrs %>% 
  filter(corr_sex_type == "between_sex" &  trait1 == trait2) %>% 
  arrange(-corr) %>% 
  my_data_table()
```




## Cuticular hydrocarbon traits {.tabset}

### Heatmap
```{r}
snp_effect_data <- load_guild("CHC") 

snp_effect_data %>% 
  make_corr_matrix() %>% make_heatmap()
```

### Fitness-trait correlations
```{r}
all_corrs <- snp_effect_data %>% make_correlation_table()

all_corrs %>% 
  filter(str_detect(trait1, "fitness") | str_detect(trait2, "fitness")) %>% 
  my_data_table()
```

### Between-sex correlations for shared traits

```{r}
all_corrs %>% 
  filter(corr_sex_type == "between_sex" &  trait1 == trait2) %>% 
  arrange(-corr) %>% 
  my_data_table()
```


## Metabolome traits {.tabset}

### Heatmap
```{r}
snp_effect_data <- load_guild("Metabolome") 

snp_effect_data %>% 
  make_corr_matrix() %>% make_heatmap()
```

### Fitness-trait correlations
```{r}
all_corrs <- snp_effect_data %>% make_correlation_table()

all_corrs %>% 
  filter(str_detect(trait1, "fitness") | str_detect(trait2, "fitness")) %>% 
  my_data_table()
```

### Between-sex correlations for shared traits

```{r}
all_corrs %>% 
  filter(corr_sex_type == "between_sex" &  trait1 == trait2) %>% 
  arrange(-corr) %>% 
  my_data_table()
```

## Sensory traits {.tabset}

### Heatmap
```{r}
snp_effect_data <- load_guild("Sensory") 

snp_effect_data %>% 
  make_corr_matrix() %>% make_heatmap()
```

### Fitness-trait correlations
```{r}
all_corrs <- snp_effect_data %>% make_correlation_table()

all_corrs %>% 
  filter(str_detect(trait1, "fitness") | str_detect(trait2, "fitness")) %>% 
  my_data_table()
```

### Between-sex correlations for shared traits

```{r}
all_corrs %>% 
  filter(corr_sex_type == "between_sex" &  trait1 == trait2) %>% 
  arrange(-corr) %>% 
  my_data_table()
```
